---
title: "cell type analysis"
output: html_notebook: default

  html_document: default
---

# Preparation

## File setup

```{r}
# Set the SCRNA_ROOT environment variable to the directory that holds the sibling
# folders CelltypeAnalysis/, SAMapH5ad/, Rsource/, SpatialMappingofAlignedCelltypes/.
# Default = this notebook's parent directory (matches the repo layout when knitting),
# so no machine-specific absolute path is stored in the file.
project.root   <- Sys.getenv("SCRNA_ROOT", unset = normalizePath(".."))
work.folder    <- file.path(project.root, "CelltypeAnalysis", "")               # trailing "" -> "/"
data.folder    <- file.path(project.root, "SAMapH5ad", "")
source.folder  <- file.path(project.root, "Rsource", "")
spatial.folder <- file.path(project.root, "SpatialMappingofAlignedCelltypes", "")
```

## Install specialty libraries
Installation only run at the first time intalled packages: limma: DE analysis scrattch.hicat: Allen institute hicat for cell type clustering Rphenograph: dependency for hicat for clustering. WGCNA: weighted gene network analysis, not used in most cases as it is not scalable for large data qlcMatrix: matrix operation
```{r}
#if (!require("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("limma")
#devtools::install_github("JinmiaoChenLab/Rphenograph")
#BiocManager::install("WGCNA")
#devtools::install_github("cysouw/qlcMatrix")
#devtools::install_github("AllenInstitute/scrattch.hicat")
```

## Load packages

```{r, echo=T, results='hide'}
suppressPackageStartupMessages({
  library(parallel) # parallelization
  library(dendextend) #dendrogram
  library(dplyr) # %in%
  library(matrixStats) #matrix operation
  library(Matrix) #matrix operation
  library(Seurat) # Seurat single cell tool kit.
  library(ggplot2) # ggplot
  library(RColorBrewer) #for making color palette
  library(stringr)#change gene name
  library(pheatmap)#generate heatmap
  library(scCustomize)
  library(SeuratDisk)#ondisk analysis
  library(harmony)
  library(tidyr)
  library(tibble)
  library(hrbrthemes)
  library(future)
  library(ggExtra)
  library(qs2)
  library(Polychrome)
})

options(future.globals.maxSize = 50000 * 1024^2) # need to set for seurat clustering when large data is used
#plan("multisession", workers = 40)

```

## Define functions

```{r}
# hicat.R and decontx.R are not sourced: none of their functions are used in
# this notebook (their libraries are loaded above anyway).
source(paste0(source.folder,"IO.R"))
source(paste0(source.folder,"CelltypeCharacterization.R"))
```









# Spatial analysis of single cell data from AI

```{r}
#calculate MERFISH distance
metadata.AI.merfish <- read.csv(file = paste0(spatial.folder,"medatadata_MERFISH.csv"))
metadata.AI.HY.merfish <- metadata.AI.merfish[metadata.AI.merfish$class %in% c("11 CNU-HYa GABA","12 HY GABA","13 CNU-HYa Glut","14 HY Glut","16 HY MM Glut"),]
subclass.coord<-metadata.AI.HY.merfish[,c("x_ccf","y_ccf","z_ccf","subclass")]
subclass.coord<-subclass.coord[subclass.coord$x_ccf>mean(subclass.coord$x_ccf),]
subclass.coord.center<-data.frame()
for (i in unique(subclass.coord$subclass)) {
  subclass <- i
  x.center.tmp <- median(subclass.coord[subclass.coord$subclass==i,][,1])
  y.center.tmp <- median(subclass.coord[subclass.coord$subclass==i,][,2])
  z.center.tmp <- median(subclass.coord[subclass.coord$subclass==i,][,3])
  subclass.coord.temp<-data.frame(subclass=subclass, x.centroid= x.center.tmp, y.centroid= y.center.tmp, z.centroid= z.center.tmp)
  subclass.coord.center <- rbind(subclass.coord.center, subclass.coord.temp)
}


subclass.coord.matrix<-as.matrix(subclass.coord.center[,2:4])
rownames(subclass.coord.matrix)<-subclass.coord.center[,1]
subclass.cor = cor(t(subclass.coord.matrix))
plot<-pheatmap(subclass.cor,clustering_distance_rows = "correlation",  clustering_distance_cols = "correlation", clustering_method = "ward.D2")
plot$tree_row$order
subclass.cor.new<-subclass.cor[plot$tree_row$order,plot$tree_col$order]
nrow(subclass.cor.new)
color1 <- colorRampPalette(brewer.pal(8, "Spectral"))(16)
subclass.color <- data.frame(c(colorRampPalette(c("#D53E4F", "#E35349"))(6),
            colorRampPalette(c("#F16943", "#F7874F"))(6),
            colorRampPalette(c("#FBA55D", "#FDBE6E"))(14),
            colorRampPalette(c("#FDD682", "#F7E58E"))(13),
            colorRampPalette(c("#ECEF94", "#DAF09A"))(12),
            colorRampPalette(c("#BEE5A0", "#A1D9A4"))(3),
            colorRampPalette(c("#81CCA4", "#62BEA6"))(6),
            colorRampPalette(c("#4AA3B1", "#3288BD"))(11)
))
rownames(subclass.color)<-rownames(subclass.cor.new)
region    <-c("POA","POA","BST","BST","BST","POA","MnPO","ZI","MS", "ZI", "BST", "MS", "STN", "ZI","PH","PH","POA","POA","POA","PM","PH","PH","LH","DMH","PH","POA","LH","DMH","DMH","DMH","MM","DMH","DMH","PVH","PVH","PVH","MM","DMH","SCH","BST","BST","BST","PM","BST","BST","DMH","PVH","PM","BST","LH","BST","PVH","PVH","PVH","MS","LH","LH","BST","AHN","BST","ARH","VMH","COA","VMH","ARH","AHN","BST","ARH","MEA","MEA","ARH")
region_tag<-c("085","087","074","121","075","118","116","109","111","093","110","066","135","101","138","139","124","106","086","140","137","134","130","103","141","090","112","127","125","102","143","107","095","097","098","-096","114","123","094","083","077","076","136","082","088","105","131","092","079","117","081","089","133","099","115","119","122","080","100","084","091","129","114","128","104","132","073","108","120","113","126")
subclass.coord.center<-cbind(subclass.coord.center, color = subclass.color[subclass.coord.center$subclass,])
subclass.coord.center<-cbind(subclass.coord.center, region = paste(region, region_tag))
library(plotly)
plot_ly(data=subclass.coord.center, x=~x.centroid, y=~y.centroid, z=~z.centroid, type="scatter3d", mode="markers", text=~subclass)
p <-plot_ly(data = subclass.coord.center, type = 'scatter3d', mode='markers')
for (i in seq_along(subclass.coord.center$color)){
  p <- p %>% add_trace(x=subclass.coord.center$x.centroid[i], y=subclass.coord.center$y.centroid[i], z=subclass.coord.center$z.centroid[i],
                       marker = list(color = subclass.coord.center$color[i], opacity=0.6, size = 5),
                       name = subclass.coord.center$subclass[i])
}

metadata.AI.plot<-scdata.AI@meta.data[,c("x","y","subclass")]

LUT<-data.frame(subclass=subclass.coord.center$subclass,
                color=subclass.coord.center$color)
color.temp <- LUT$color[match(metadata.AI.plot$subclass, LUT$subclass)]


metadata.AI.plot <- cbind(metadata.AI.plot, color.temp)
metadata.AI.plot$color.temp[is.na(metadata.AI.plot$color.temp)]<-"lightgrey"


ggplot(metadata.AI.plot, aes(x=x, y=y)) +
  geom_point(size=0.1, color="#A9A9A980")+
  geom_point(data=metadata.AI.plot, aes(x=x,y=y), size=0.1, color=metadata.AI.plot$color)+
  theme_void() + 
  theme(legend.position = "none") +
  coord_fixed()

ggplot(metadata.AI.plot, aes(x=x, y=y)) +
  geom_point(size=0.1, color="#A9A9A980")+
  geom_point(data=metadata.AI.plot, aes(x=x,y=y), size=0.1, color=scdata.AI@meta.data$subclass_color)+
  theme_void() + 
  theme(legend.position = "none") +
  coord_fixed()

ggplot(metadata.AI.plot, aes(x=x, y=y)) +
  geom_point(size=0.1, color="#A9A9A980")+
  geom_point(data=metadata.AI.plot, aes(x=x,y=y), size=0.1, color=scdata.AI@meta.data$class_color)+
  theme_void() + 
  theme(legend.position = "none") +
  coord_fixed()

#analyze supertype level

supertype.coord<-metadata.AI.HY.merfish[,c("x_ccf","y_ccf","z_ccf","supertype")]
supertype.coord<-supertype.coord[supertype.coord$x_ccf>mean(supertype.coord$x_ccf),]
supertype.coord.center<-data.frame()
for (i in unique(supertype.coord$supertype)) {
  supertype <- i
  x.center.tmp <- median(supertype.coord[supertype.coord$supertype==i,][,1])
  y.center.tmp <- median(supertype.coord[supertype.coord$supertype==i,][,2])
  z.center.tmp <- median(supertype.coord[supertype.coord$supertype==i,][,3])
  supertype.coord.temp<-data.frame(supertype=supertype, x.centroid= x.center.tmp, y.centroid= y.center.tmp, z.centroid= z.center.tmp)
  supertype.coord.center <- rbind(supertype.coord.center, supertype.coord.temp)
}


supertype.coord.matrix<-as.matrix(supertype.coord.center[,2:4])
rownames(supertype.coord.matrix)<-supertype.coord.center[,1]
supertype.cor = cor(t(supertype.coord.matrix))
plot<-pheatmap(supertype.cor,clustering_distance_rows = "correlation",  clustering_distance_cols = "correlation", clustering_method = "ward.D2", cutree_rows = 8)
plot$tree_row$order
supertype.cor.new<-supertype.cor[plot$tree_row$order,plot$tree_col$order]
nrow(supertype.cor.new)
color1 <- brewer.pal(8, "Spectral")
supertype.coord.center<-cbind(supertype.coord.center,cutree=cutree(plot$tree_row, k = 8))
LUT <- data.frame(value = 1:8, color = color1)
supertype.color <- LUT$color[match(supertype.coord.center$cutree, LUT$value)]


supertype.coord.center<-cbind(supertype.coord.center, color = supertype.color)
plot_ly(data=supertype.coord.center, x=~x.centroid, y=~y.centroid, z=~z.centroid, type="scatter3d", mode="markers", text=~supertype)
p <-plot_ly(data = supertype.coord.center, type = 'scatter3d', mode='markers')
for (i in seq_along(supertype.coord.center$color)){
  p <- p %>% add_trace(x=supertype.coord.center$x.centroid[i], y=supertype.coord.center$y.centroid[i], z=supertype.coord.center$z.centroid[i],
                       marker = list(color = supertype.coord.center$color[i], opacity=0.6, size = 5),
                       name = supertype.coord.center$subclass[i])
}

metadata.AI.plot<-scdata.AI@meta.data[,c("x","y","supertype")]

LUT<-data.frame(supertype=supertype.coord.center$supertype,
                color=supertype.coord.center$color)
color.temp <- LUT$color[match(metadata.AI.plot$supertype, LUT$supertype)]


metadata.AI.plot <- cbind(metadata.AI.plot, color.temp)
metadata.AI.plot$color.temp[is.na(metadata.AI.plot$color.temp)]<-"lightgrey"


ggplot(metadata.AI.plot, aes(x=x, y=y)) +
  geom_point(size=0.1, color="#A9A9A980")+
  geom_point(data=metadata.AI.plot, aes(x=x,y=y), size=0.1, color=metadata.AI.plot$color)+
  theme_void() + 
  theme(legend.position = "none") +
  coord_fixed()

ggplot(metadata.AI.plot, aes(x=x, y=y)) +
  geom_point(size=0.1, color="#A9A9A980")+
  geom_point(data=metadata.AI.plot, aes(x=x,y=y), size=0.1, color=scdata.AI@meta.data$supertype_color)+
  theme_void() + 
  theme(legend.position = "none") +
  coord_fixed()


#Calculate distance and output to downstream analysis
nrow(subclass.coord.center)
rownames(subclass.coord.center)<-subclass.coord.center$subclass
subclass.coord.center<-subclass.coord.center[,2:4]
colnames(subclass.coord.center)<-c("x","y","z")
write.csv(subclass.coord.center, file="AI.subclass.spatialcentroid.csv")
```





# Import data




## Import Allen insitute hypothalamus data
```{r}
count<-readMM(file = paste(data.folder,"AI_HY/AI_HY.mtx",sep=""))
metadata.AI <- read.csv(file = paste(data.folder,"AI_HY/AI_HY_metadata.csv",sep=""))
rownames(metadata.AI) <- metadata.AI$cell_label
saveRDS(metadata.AI, file = paste(work.folder,"metadata.AI.rds",sep=""))
metadata.AI <- readRDS(file = paste(work.folder,"metadata.AI.rds",sep=""))


ggplot(metadata.AI, aes(x=x, y=y)) +
  geom_point(size=0.1, color="#A9A9A980")+
  geom_point(data=metadata.AI, aes(x=x,y=y), size=0.1, color=metadata.AI$subclass_color)+
  theme_void() + 
  theme(legend.position = "none") +
  coord_fixed()


cellID <- read.csv(file = paste(data.folder, "AI_HY/AI_HY_cellID.csv", sep=""), header=T, row.names = 1)
geneID <- read.csv(file = paste(data.folder, "AI_HY/AI_HY_geneID.csv", sep=""), header=T, row.names = 1)

#load metadata
rownames(count) <- cellID[,1]
colnames(count) <- geneID[,1]
count<-t(count)
scdata.AI <- CreateSeuratObject(counts = count, assay = "RNA", meta.data = metadata.AI, project = "Mmusculus")
scdata.AI <- DietSeurat(scdata.AI)

Idents(scdata.AI)<-"class"
scdata.AI<-subset(x=scdata.AI, idents = c("11 CNU-HYa GABA","12 HY GABA","13 CNU-HYa Glut","14 HY Glut","16 HY MM Glut" ))

saveRDS(scdata.AI, file = "scdata.AI.HY.rds")
scdata.AI<-readRDS(file="scdata.AI.HY.rds")




Idents(scdata.AI)<-"subclass"
scdata.AI <- subset(x = scdata.AI, subset = nFeature_RNA > 0 )
scdata.AI <- subset(x = scdata.AI, subset = nCount_RNA > 0 )
scdata.AI.sub1000 <- subset(scdata.AI, downsample=1000)
VlnPlot(scdata.AI.sub1000, features=c("nCount_RNA","nFeature_RNA"))
scdata.AI.sub1000 <- SCTransform(scdata.AI.sub1000, vst.flavor = "v2", verbose = FALSE) 


ggplot(scdata.AI.sub1000@meta.data, aes(x=x, y=y)) +
  geom_point(size=0.1, color="#A9A9A980")+
  geom_point(data=scdata.AI.sub1000@meta.data, aes(x=x,y=y), size=0.1, color=scdata.AI.sub1000@meta.data$subclass_color)+
  theme_void() + 
  theme(legend.position = "none") +
  coord_fixed()



saveRDS(scdata.AI.sub1000, file = "scdata.AI.sub1000.rds")
scdata.AI.sub1000<-readRDS(file = "scdata.AI.sub1000.rds")
```


## Import Prairie vole data

```{r}
# import data to Seurat format
scdata.MO <- ImportH5ad(folder = paste0(data.folder, "MO_dat_07262026_yw/"), org = "MO")
# default processing
scdata.MO <- SeuratProcess(scdata.MO, "key",ncore = 40, run.iterative = FALSE)

#plot
n_cls <- length(unique(scdata.MO@meta.data$eq_subclass))
# Dynamically generate N > 256 visually distinct colors in HCL space
seed_cols <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF")
color <- as.vector(createPalette(n_cls, seed_cols, M = max(1000, n_cls * 3)))
p <- DimPlot(scdata.MO, reduction = "tsne", group.by = "eq_subclass", cols = color) + NoLegend()


qs_save(scdata.MO, file = "scdata.MO.qs2", nthreads = 24)
#scdata.MO <- qs_read("scdata.MO.qs2", nthreads = 24)
```




## Import Japanese quail data
```{r}
# import data to Seurat format
scdata.CJ <- ImportH5ad(folder = paste0(data.folder, "CJ_dat_07262026_yw/"), org = "CJ")
# default processing
scdata.CJ <- SeuratProcess(scdata.CJ, "key", ncore = 40, run.iterative = FALSE)
#color <- rep(brewer.pal(9, "Set1"), length(unique(scdata.CJ@meta.data$eq_subclass))/9+1)
#DimPlot(scdata.CJ, reduction="tsne", group.by = "eq_subclass", cols=color)+NoLegend()

# save data
qs_save(scdata.CJ, file = "scdata.CJ.qs2", nthreads = 24)
#scdata.CJ <- qs_read("scdata.CJ.qs2", nthreads = 24)
```




## Import green anole data
```{r}
# import data to Seurat format
scdata.AC <- ImportH5ad(folder = paste0(data.folder, "AC_dat_07262026_yw/"), org = "AC")
# default processing
scdata.AC <- SeuratProcess(scdata.AC, "key", ncore = 40, run.iterative = FALSE)
#color <- rep(brewer.pal(9, "Set1"), length(unique(scdata.AC@meta.data$eq_subclass))/9+1)
#DimPlot(scdata.AC, reduction="tsne", group.by = "eq_subclass", cols=color)+NoLegend()
#p2<-DimPlot(scdata.AC, reduction="pca", group.by = "eq_subclass", cols=color)+NoLegend()+xlim(-150,20)+ylim(-120, 20)


# save data
qs_save(scdata.AC, file = "scdata.AC.qs2", nthreads = 24)
#scdata.AC <- qs_read("scdata.AC.qs2", nthreads = 24)
```

## Import xenopus data
```{r}
# import data to Seurat format
scdata.XT <- ImportH5ad(folder = paste0(data.folder, "XT_dat_07262026_yw/"), org = "XT")

# default processing
scdata.XT <- SeuratProcess(scdata.XT, "key", ncore = 40, run.iterative = FALSE)
#color <- rep(brewer.pal(9, "Set1"), length(unique(scdata.XT@meta.data$eq_subclass))/9+1)
#DimPlot(scdata.XT, reduction="tsne", group.by = "eq_subclass", cols=color)+NoLegend()

# save data
qs_save(scdata.XT, file = "scdata.XT.qs2", nthreads = 24)
#scdata.XT <- qs_read("scdata.XT.qs2", nthreads = 24)
```

```{r}
scdata.XT.old <- ImportH5ad(folder = paste0(data.folder, "XT_dat_nounlabeled_06232026_yw/"), org = "XT")
# default processing
scdata.XT.old <- SeuratProcess(scdata.XT.old, "key", ncore = 40, run.iterative = FALSE)
```


## Import zebrafish data
```{r}
# import data to Seurat format
scdata.DR <- ImportH5ad(folder = paste0(data.folder, "DR_dat_07262026_yw/"), org = "DR")
# default processing
scdata.DR <- SeuratProcess(scdata.DR, "key", ncore = 40, run.iterative = FALSE)

# save data
qs_save(scdata.DR, file = "scdata.DR.qs2", nthreads = 24)
#scdata.DR <- qs_read("scdata.DR.qs2", nthreads = 24)
```








Neurotransmitter analysis

AI criteria

Glutamatergic (Glut): Slc17a6 (also known as Vglut2), Slc17a7 (Vglut1) or Slc17a8 (Vglut3).

GABAergic (GABA): (Slc32a1 (Vgat) or Slc18a2 (Vmat2)) and (Gad1, Gad2 or Aldh1a1).

Glycinergic (Glyc): Slc6a5.

Cholinergic (Chol): Slc18a3 (Vacht) and Chat.

Dopaminergic (Dopa): (Slc6a3 (Dat) or Slc18a2) and (Th and Ddc).

Serotonergic (Sero): (Slc6a4 (Sert) or Slc18a2) and (Tph2 and Ddc).

Noradrenergic (Nora): (Slc6a2 (Net) or Slc18a2) and Dbh.

Histaminergic (Hist): Slc18a2 and Hdc.



# Neurotransmitter type annotation

Allen institute criteria

Glutamatergic (Glut): Slc17a6 (also known as Vglut2), Slc17a7 (Vglut1) or Slc17a8 (Vglut3).
GABAergic (GABA): (Slc32a1 (Vgat) or Slc18a2 (Vmat2)) and (Gad1, Gad2 or Aldh1a1).
Glycinergic (Glyc): Slc6a5.
Cholinergic (Chol): Slc18a3 (Vacht) and Chat.
Dopaminergic (Dopa): (Slc6a3 (Dat) or Slc18a2) and (Th and Ddc).
Serotonergic (Sero): (Slc6a4 (Sert) or Slc18a2) and (Tph2 and Ddc).
Noradrenergic (Nora): (Slc6a2 (Net) or Slc18a2) and Dbh.
Histaminergic (Hist): Slc18a2 and Hdc.
We used a stringent expression threshold of log2(CPM)>3 for these genes to assign neurotransmitter identity to each cluster.


Idents, e.g. "eq_subclass" should be changed to the col name of the equivalent subclass.

## function to calculate per cell type average expression 

This function may generate wrong estimate if cells within one cluster is < 30
set seed for reproducibility
```{r}
  NT.marker.threshold.fast <- function(NT.marker, scdata, cluster.label, downsample = 500, seed = 61) {
      
      # Set seed for reproducible downsampling
      if (!is.null(seed)) {
        set.seed(seed)
      }
      
      meta_labels <- scdata[[cluster.label]][, 1]
      cell_names <- rownames(scdata@meta.data)
      unique_clusters <- unique(meta_labels)
      
      # 1. Fast cell sampling without deep-copying Seurat object
      if (!is.null(downsample)) {
        cells_to_keep <- unlist(lapply(unique_clusters, function(cluster) {
          cluster_cells <- cell_names[meta_labels == cluster]
          if (length(cluster_cells) <= downsample) {
            return(cluster_cells)
          } else {
            return(sample(cluster_cells, downsample, replace = FALSE))
          }
        }), use.names = FALSE)
      } else {
        cells_to_keep <- cell_names
      }
      
      # 2. Extract sparse counts for target gene & library sizes directly
      counts_mat <- GetAssayData(scdata, assay = "RNA", layer = "counts")
      if (!NT.marker %in% rownames(counts_mat)) {
        stop(paste("Error: The gene", NT.marker, "was not found in the RNA assay."))
      }
      
      gene_counts <- counts_mat[NT.marker, cells_to_keep]
      total_counts <- scdata$nCount_RNA[cells_to_keep]
      sub_meta <- meta_labels[match(cells_to_keep, cell_names)]
      
      # 3. Sparse-Aware ln(CPM + 1) — Calculate ONLY on non-zero entries
      ln_cpm <- numeric(length(gene_counts))
      nz_idx <- which(gene_counts > 0)
      if (length(nz_idx) > 0) {
        ln_cpm[nz_idx] <- log(1e6 * gene_counts[nz_idx] / total_counts[nz_idx] + 1)
      }
      
      # 4. Vectorized one-pass aggregation (no cluster loop required)
      sub_meta_fac <- factor(sub_meta, levels = unique_clusters)
      mean_lnCPM <- tapply(ln_cpm, sub_meta_fac, mean)
      pct_expr <- tapply(gene_counts > 0, sub_meta_fac, function(x) mean(x) * 100)
      
      marker.exp <- data.frame(
        celltype = unique_clusters,
        mean_lnCPM = as.numeric(mean_lnCPM),
        pct.expr = as.numeric(pct_expr),
        row.names = unique_clusters
      )
      
      return(marker.exp)
    }

```


## Annotate Xenopus tropicalis NT types


```{r}

#initiate a table to label NT types
Idents(scdata.XT)<-"eq_subclass"
plot.chart <- DotPlot(scdata.XT,c("slc32a1","slc17a6","chat","th","tph2"),cols = c("lightgrey","darkred"))
XT.nt <- data.frame(scdata.XT@meta.data[,c("eq_subclass")])
colnames(XT.nt)<-"eq_subclass"
XT.nt <- XT.nt[duplicated(XT.nt$eq_subclass)==F, ]
XT.nt <- data.frame(subclass=XT.nt, Glut=rep(0, length(XT.nt)), Gaba=rep(0, length(XT.nt)), Dopa=rep(0, length(XT.nt)), Chol=rep(0, length(XT.nt)), Sert=rep(0, length(XT.nt)), Nora=rep(0, length(XT.nt)), Gly=rep(0, length(XT.nt)), Hist=rep(0, length(XT.nt)), Neurotransmitter=rep(0, length(XT.nt)))



#find an arbituary threshold for neurotransmitter markers.




#Dopaminergic
#ddc not expressed.
plot.dopa.xt <- DotPlot(scdata.XT,c("slc6a3","SLC18A2","th"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
th.exp <- NT.marker.threshold("th", scdata.XT, cluster.label="eq_subclass")
slc18a2.exp <- NT.marker.threshold("SLC18A2", scdata.XT, "eq_subclass")
dat.exp <- NT.marker.threshold("slc6a3", scdata.XT, "eq_subclass")

slc18a2 <- slc18a2.exp[slc18a2.exp[,2]>1.5,1]
dat <- dat.exp[dat.exp[,2]>1,1]
th <- th.exp[th.exp[,2]>2,1]

dopa <- intersect(union(slc18a2.exp[slc18a2.exp[,2]>1.5,1], dat.exp[dat.exp[,2]>1,1]), th.exp[th.exp[,2]>2,1])
#dopa thresold 
XT.nt$Dopa[XT.nt$subclass %in% dopa]<-1

#Gabaergic
#gad1 has 2 paralogs
plot.gaba.xt <-DotPlot(scdata.XT,c("slc32a1","SLC18A2","gad1.1", "gad1.2","gad2","aldh1a1"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
DotPlot(scdata.XT,c("gad1.2"),cols = c("lightgrey","darkred"), assay = "RNA")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
DotPlot(scdata.XT,c("aldh1a1"),cols = c("lightgrey","darkred"), assay = "RNA")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
slc32a1.exp <- NT.marker.threshold("slc32a1", scdata.XT, "eq_subclass")
gad1.1.exp <- NT.marker.threshold("gad1.1", scdata.XT, "eq_subclass")
gad1.2.exp <- NT.marker.threshold("gad1.2", scdata.XT, "eq_subclass")
gad2.exp <- NT.marker.threshold("gad2", scdata.XT, "eq_subclass")
aldh1a1.exp <- NT.marker.threshold("aldh1a1", scdata.XT, "eq_subclass")
slc17a6.exp <- NT.marker.threshold("slc17a6", scdata.XT, "eq_subclass")

gaba.all.exp<-data.frame(celltype=slc32a1.exp$celltype, slc32a1=slc32a1.exp$mean_lnCPM, gad1.1=gad1.1.exp$mean_lnCPM, gad1.2=gad1.2.exp$mean_lnCPM, gad2=gad2.exp$mean_lnCPM, slc17a6=slc17a6.exp$mean_lnCPM)
# Save the scatter plot in a variable
p1 <- ggplot(gaba.all.exp, aes(x = slc17a6, y = slc32a1)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p1, type = "histogram")

slc32a1 <- slc32a1.exp[slc32a1.exp[,2]>1.3,1]
p2 <- ggplot(gaba.all.exp, aes(x = slc17a6, y = gad1.1)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p2, type = "histogram")

gad1.1 <- gad1.1.exp[gad1.1.exp[,2]>1,1]
gad1.2 <- gad1.2.exp[gad1.2.exp[,2]>1,1]

p3 <- ggplot(gaba.all.exp, aes(x = slc17a6, y = gad2)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p3, type = "histogram")
gad2 <- gad2.exp[gad2.exp[,2]>2.5,1]


aldh1a1 <- aldh1a1.exp[aldh1a1.exp[,2]>0.5,1]

gaba <- intersect(union(slc18a2, slc32a1), union(gad1.2, union(gad1.1,union(gad2, aldh1a1))))
gaba[order(gaba)]
XT.nt$Gaba[XT.nt$subclass %in% gaba]<-1

#Glutamatergic
plot.glut.xt<-DotPlot(scdata.XT,c("slc17a6","slc17a7","slc17a8", "slc32a1"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
#DotPlot(scdata.XT,c("slc17a8"),cols = c("lightgrey","darkred"), assay = "RNA")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
slc17a6.exp <- NT.marker.threshold("slc17a6", scdata.XT, "eq_subclass")
slc17a7.exp <- NT.marker.threshold("slc17a7", scdata.XT, "eq_subclass")
slc17a8.exp <- NT.marker.threshold("slc17a8", scdata.XT, "eq_subclass")

glut.all.exp<-data.frame(celltype=slc32a1.exp$celltype, slc32a1=slc32a1.exp$mean_lnCPM, slc17a6=slc17a6.exp$mean_lnCPM, slc17a7=slc17a7.exp$mean_lnCPM, slc17a8=slc17a8.exp$mean_lnCPM)
p2 <- ggplot(glut.all.exp, aes(x = slc17a7, y = slc32a1)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p2, type = "histogram")

slc17a6 <- slc17a6.exp[slc17a6.exp[,2]>2,1]
slc17a7 <- slc17a7.exp[slc17a7.exp[,2]>2,1]
slc17a8 <- slc17a8.exp[slc17a8.exp[,2]>2,1]
glut <- union(slc17a6, union(slc17a7, slc17a8))
glut[order(glut)]
XT.nt$Glut[XT.nt$subclass %in% glut]<-1



#Serotoninergic
#Serotonergic (Sero): (Slc6a4 (Sert) or Slc18a2) and (Tph2 and Ddc).
#Tph1: ENSXETG00000048806
#Ddc not detected.
plot.sero.xt<-DotPlot(scdata.XT,c("slc6a4","SLC18A2","tph2","ENSXETG00000048806"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
tph2.exp <- NT.marker.threshold("tph2", scdata.XT, "eq_subclass")
tph1.exp <- NT.marker.threshold("ENSXETG00000048806", scdata.XT, "eq_subclass")
slc6a4.exp <- NT.marker.threshold("slc6a4", scdata.XT, "eq_subclass")
slc6a4 <- slc6a4.exp[slc6a4.exp[,2]>3,1]
tph2 <- tph2.exp[tph2.exp[,2]>3,1]
tph1 <- tph1.exp[tph1.exp[,2]>3,1]
sert <- intersect(union(slc6a4, slc18a2), union(tph2,tph1))
XT.nt$Sert[XT.nt$subclass %in% sert]<-1

#Glycinergic
#Glycinergic (Glyc): Slc6a5.
plot.gly.xt<-DotPlot(scdata.XT,c("slc6a5"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc6a5.exp <- NT.marker.threshold("slc6a5", scdata.XT, "eq_subclass")
gly <- slc6a5.exp[slc6a5.exp[,2]>0.4,1]
XT.nt$Gly[XT.nt$subclass %in% gly]<-1

#Noradrenergic
#Noradrenergic (Nora): (Slc6a2 (Net) or Slc18a2) and Dbh.
plot.nora.xt<-DotPlot(scdata.XT,c("slc6a2","SLC18A2","dbh"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
#DotPlot(scdata.XT,c("slc6a2","dbh"),cols = c("lightgrey","darkred"), assay = "RNA")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
slc6a2.exp <- NT.marker.threshold("slc6a2", scdata.XT, "eq_subclass")
dbh.exp <- NT.marker.threshold("dbh", scdata.XT, "eq_subclass")
nora <- intersect(slc6a2.exp[slc6a2.exp[,2]>1,1], dbh.exp[dbh.exp[,2]>1,1])
XT.nt$Nora[XT.nt$subclass %in% nora]<-1


#Cholinergic
#Cholinergic (Chol): Slc18a3 (Vacht) and Chat.
plot.chol.xt<-DotPlot(scdata.XT,c("slc18a3","chat"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc18a3.exp <- NT.marker.threshold("slc18a3", scdata.XT, "eq_subclass")
chat.exp <- NT.marker.threshold("chat", scdata.XT, "eq_subclass")
chol <- intersect(slc18a3.exp[slc18a3.exp[,2]>1,1], chat.exp[chat.exp[,2]>1,1])
XT.nt$Chol[XT.nt$subclass %in% chol]<-1


#histaminergic
#Histaminergic (Hist): Slc18a2 and Hdc.
plot.hist.xt<-DotPlot(scdata.XT,c("SLC18A2","hdc"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
hdc.exp <- NT.marker.threshold("hdc", scdata.XT, "eq_subclass")
hist <- intersect(slc18a2.exp[slc18a2.exp[,2]>1.5,1], hdc.exp[hdc.exp[,2]>1,1])
XT.nt$Hist[XT.nt$subclass %in% hist]<-1

NT <- c("Glut","Gaba", "Dopa", "Chol", "Sert", "Nora", "Gly", "Hist")
NT.1 <- t(replicate(nrow(XT.nt), NT))
NT.2 <- t(apply(XT.nt[,2:9], 1, as.logical))
for (i in 1:nrow(XT.nt)) {
  XT.nt$Neurotransmitter[i] <- paste(NT.1[i,][NT.2[i,]], collapse="-")
}
XT.nt$Neurotransmitter[XT.nt$Neurotransmitter==""]<-"unclear"
#save data
write.csv(XT.nt, file="./XT.nt.type.csv")
XT.nt<-read.csv(file="./XT.nt.type.csv")

#add to metadata
scdata.XT@meta.data$Neurotransmitter <- XT.nt$Neurotransmitter[match(scdata.XT@meta.data$eq_subclass, XT.nt$subclass)]
color <- colorRampPalette(brewer.pal(9, "Set1")[1:8])(12)
color <- c(color,"grey")

color=c("#FF3358","#73E785","#fcf04b","#fcf04b","#820e57","#ff7621","#2B93DF","#fcf04b","#0a9964","#820e57","#4F2E91FF","#99999960")
plot.xt.label<-DimPlot(scdata.XT, group.by = "Neurotransmitter", cols = color, reduction = "tsne")+coord_fixed()
qs_save(scdata.XT, file = "scdata.XT.qs2", nthreads = 24)
# Read it back in using 24 CPU cores
scdata.XT <- qs_read("scdata.XT.qs2", nthreads = 24)

```






## Annotate Coturnix Japonica NT type


```{r}
Idents(scdata.CJ)<-"eq_subclass"
CJ.nt <- data.frame(scdata.CJ@meta.data[,c("eq_subclass")])
colnames(CJ.nt)<-"eq_subclass"
CJ.nt <- CJ.nt[duplicated(CJ.nt$eq_subclass)==F, ]
CJ.nt <- data.frame(subclass=CJ.nt, Glut=rep(0, length(CJ.nt)), Gaba=rep(0, length(CJ.nt)), Dopa=rep(0, length(CJ.nt)), Chol=rep(0, length(CJ.nt)), Sert=rep(0, length(CJ.nt)), Nora=rep(0, length(CJ.nt)), Gly=rep(0, length(CJ.nt)), Hist=rep(0, length(CJ.nt)), Neurotransmitter=rep(0, length(CJ.nt)))

#find an arbituary threshold for neurotransmitter markers.

#Dopaminergic
#Dopaminergic (Dopa): (Slc6a3 (Dat) or Slc18a2) and (Th and Ddc).
#DAT not found in annotation. SLC6A2 has a 95% BLASTP match. The annotated Dopa cell types all have the expression of this gene. However, not all SLC6a2 expression cells are annotated as Dopa because of the lack of Th and Ddc.
CJ.NT.plot.Dopa <- DotPlot(scdata.CJ,c("SLC6A2","SLC18A2", "DDC","TH"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc6a2.exp <- NT.marker.threshold("SLC6A2", scdata.CJ, "eq_subclass")
ddc.exp <- NT.marker.threshold("DDC", scdata.CJ, "eq_subclass")
th.exp <- NT.marker.threshold("TH", scdata.CJ, "eq_subclass")
slc18a2.exp <- NT.marker.threshold("SLC18A2", scdata.CJ, "eq_subclass")
slc6a2 <- slc6a2.exp[slc6a2.exp[,2]>0.4,1]
slc18a2 <- slc18a2.exp[slc18a2.exp[,2]>1.5,1]
th <- th.exp[th.exp[,2]>1,1]
ddc <- ddc.exp[ddc.exp[,2]>1,1]
dopa <- intersect(union(slc6a2,slc18a2), intersect(ddc, th))
#dopa thresold th>3, dat>3
CJ.nt$Dopa[CJ.nt$subclass %in% dopa]<-1

#Gabaergic
#GABAergic (GABA): (Slc32a1 (Vgat) or Slc18a2 (Vmat2)) and (Gad1, Gad2 or Aldh1a1).
CJ.NT.plot.Gaba <- DotPlot(scdata.CJ,c("SLC32A1","SLC18A2","GAD1", "GAD2","SLC17A6","ALDH1A1"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc32a1.exp <- NT.marker.threshold("SLC32A1", scdata.CJ, "eq_subclass")
gad1.exp <- NT.marker.threshold("GAD1", scdata.CJ, "eq_subclass")
gad2.exp <- NT.marker.threshold("GAD2", scdata.CJ, "eq_subclass")
aldh1a1.exp <- NT.marker.threshold("ALDH1A1", scdata.CJ, "eq_subclass")
slc17a6.exp <- NT.marker.threshold("SLC17A6", scdata.CJ, "eq_subclass")
slc17a7.exp <- NT.marker.threshold("ENSCJPG00005000657", scdata.CJ, "eq_subclass")
slc17a8.exp <- NT.marker.threshold("SLC17A8", scdata.CJ, "eq_subclass")



gaba.all.exp<-data.frame(celltype=slc32a1.exp$celltype, slc32a1=slc32a1.exp$mean_lnCPM, gad1=gad1.exp$mean_lnCPM, gad2=gad2.exp$mean_lnCPM, slc17a6=slc17a6.exp$mean_lnCPM, slc17a7=slc17a7.exp$mean_lnCPM, slc17a8=slc17a8.exp$mean_lnCPM)
# Save the scatter plot in a variable
p1 <- ggplot(gaba.all.exp, aes(x = slc17a6, y = slc32a1)) +
  geom_point()
p1 <- ggplot(gaba.all.exp, aes(x = slc17a7, y = slc32a1)) +
  geom_point()
p1 <- ggplot(gaba.all.exp, aes(x = slc17a8, y = slc32a1)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p1, type = "histogram")

slc32a1 <- slc32a1.exp[slc32a1.exp[,2]>0.5,1]
p2 <- ggplot(gaba.all.exp, aes(x = slc17a6, y = gad1)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p2, type = "histogram")

gad1 <- gad1.exp[gad1.exp[,2]>1.5,1]


p3 <- ggplot(gaba.all.exp, aes(x = slc17a6, y = gad2)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p3, type = "histogram")
gad2 <- gad2.exp[gad2.exp[,2]>3,1]


gaba <- intersect(union(slc18a2, slc32a1), union(gad2, union(gad1,union(gad2, aldh1a1))))
gaba[order(gaba)]
CJ.nt$Gaba[CJ.nt$subclass %in% gaba]<-1

#Glutamatergic
#Glutamatergic (Glut): Slc17a6 (also known as Vglut2), Slc17a7 (Vglut1) or Slc17a8 (Vglut3).


slc17a6 <- slc17a6.exp[slc17a6.exp[,2]>2,1]
slc17a7 <- slc17a7.exp[slc17a7.exp[,2]>1,1]
slc17a8 <- slc17a8.exp[slc17a8.exp[,2]>1,1]
glut <- union(slc17a7, union(slc17a6, slc17a8))
CJ.nt$Glut[CJ.nt$subclass %in% glut]<-1


#Serotoninergic
#Serotonergic (Sero): (Slc6a4 (Sert) or Slc18a2) and (Tph2 or Tph1) and Ddc).
CJ.NT.plot.Sert <- DotPlot(scdata.CJ,c("SLC6A4","SLC18A2","TPH2", "TPH1", "DDC"),cols = c("lightgrey","darkred"), scale.max = 20)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
tph2.exp <- NT.marker.threshold("TPH2", scdata.CJ, "eq_subclass")
slc6a4.exp <- NT.marker.threshold("SLC6A4", scdata.CJ, "eq_subclass")
ddc.exp <- NT.marker.threshold("DDC", scdata.CJ, "eq_subclass")
tph1.exp <- NT.marker.threshold("TPH1", scdata.CJ, "eq_subclass")

sert.all.exp<-data.frame(celltype=tph2.exp$celltype, tph2=tph2.exp$mean_lnCPM, slc6a4=slc6a4.exp$mean_lnCPM, ddc=ddc.exp$mean_lnCPM, tph1=tph1.exp$mean_lnCPM)
# Save the scatter plot in a variable
p1 <- ggplot(sert.all.exp, aes(x = tph2, y = slc6a4)) +
  geom_point()
p1 <- ggplot(sert.all.exp, aes(x = tph2, y = ddc)) +
  geom_point()
p1 <- ggplot(sert.all.exp, aes(x = tph2, y = tph1)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p1, type = "histogram")
# there are many serotoninergic neurons that are ddc negative.




slc6a4 <- slc6a4.exp[slc6a4.exp[,2]>0.09,1]
tph2 <- tph2.exp[tph2.exp[,2]>1,1]
tph1 <- tph1.exp[tph1.exp[,2]>1,1]
ddc <- ddc.exp[ddc.exp[,2]>1,1]
sert <- intersect(slc6a4, intersect(union(tph1,tph2), ddc))
CJ.nt$Sert[CJ.nt$subclass %in% sert]<-1
#no serotoninergic neurons in quail hypothalamus

#Glycinergic
#Glycinergic (Glyc): Slc6a5.
CJ.NT.plot.Gly <- DotPlot(scdata.CJ,c("SLC6A5"),cols = c("lightgrey","darkred"),scale.max = 20)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc6a5.exp <- NT.marker.threshold("SLC6A5", scdata.CJ, "eq_subclass")
gly <- slc6a5.exp[slc6a5.exp[,2]>0.1,1]
CJ.nt$Gly[CJ.nt$subclass %in% gly]<-1

#Noradrenergic
#Noradrenergic (Nora): (Slc6a2 (Net) or Slc18a2) and Dbh.
CJ.NT.plot.Nora <- DotPlot(scdata.CJ,c("SLC6A2","DBH"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc6a2.exp <- NT.marker.threshold("SLC6A2", scdata.CJ, "eq_subclass")
dbh.exp <- NT.marker.threshold("DBH", scdata.CJ, "eq_subclass")
slc6a2 <- slc6a2.exp[slc6a2.exp[,2]>0.4,1]
dbh <- dbh.exp[dbh.exp[,2]>1,1]
nora <- intersect(union(slc6a2,slc18a2), dbh)
CJ.nt$Nora[CJ.nt$subclass %in% nora]<-1


#Cholinergic
#Cholinergic (Chol): Slc18a3 (Vacht) and Chat.
CJ.NT.plot.Chol <- DotPlot(scdata.CJ,c("SLC18A3","CHAT"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc18a3.exp <- NT.marker.threshold("SLC18A3", scdata.CJ, "eq_subclass")
chat.exp <- NT.marker.threshold("CHAT", scdata.CJ, "eq_subclass")
slc18a3 <- slc18a3.exp[slc18a3.exp[,2]>1,1]
chat <- chat.exp[chat.exp[,2]>3,1]
chol <- intersect(slc18a3, chat)
CJ.nt$Chol[CJ.nt$subclass %in% chol]<-1


#histaminergic
#Histaminergic (Hist): Slc18a2 and Hdc.
CJ.NT.plot.Hist <- DotPlot(scdata.CJ,c("SLC18A2","HDC"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
hdc.exp <- NT.marker.threshold("HDC", scdata.CJ, "eq_subclass")
hdc <- hdc.exp[hdc.exp[,2]>1,1]
hist <- intersect(slc18a2, hdc)
CJ.nt$Hist[CJ.nt$subclass %in% hist]<-1

NT <- c("Glut","Gaba", "Dopa", "Chol", "Sert", "Nora", "Gly", "Hist")
NT.1 <- t(replicate(nrow(CJ.nt), NT))
NT.2 <- t(apply(CJ.nt[,2:9], 1, as.logical))
for (i in 1:nrow(CJ.nt)) {
  CJ.nt$Neurotransmitter[i] <- paste(NT.1[i,][NT.2[i,]], collapse="-")
}
CJ.nt$Neurotransmitter[CJ.nt$Neurotransmitter==""]<-"unclear"
#save data
write.csv(CJ.nt, file="./CJ.nt.type.csv")
CJ.nt<-read.csv(file="./CJ.nt.type.csv")
#add to metadata
scdata.CJ@meta.data$neurotransmitter <- CJ.nt$Neurotransmitter[match(scdata.CJ@meta.data$eq_subclass, CJ.nt$subclass)]
color <- colorRampPalette(brewer.pal(9, "Set1")[1:8])(13)
color <- c("grey",color)
DimPlot(scdata.CJ, group.by = "neurotransmitter", cols = color, reduction = "tsne")+coord_fixed()
aggregate(rep(1, nrow(scdata.CJ@meta.data)), by = list(x = scdata.CJ@meta.data$neurotransmitter, y = scdata.CJ@meta.data$eq_subclass), sum)

#future(saveRDS(scdata.CJ, file="scdata.CJ.rds"))
qs_save(scdata.CJ, file = "scdata.CJ.qs2", nthreads = 8)
# Read it back in using 4 CPU cores
scdata.CJ <- qs_read("scdata.CJ.qs2", nthreads = 8)
View(CJ.nt)
```



## Annotate Microtus ochrogaster NT type

```{r}
Idents(scdata.MO)<-"eq_subclass"
MO.nt <- data.frame(scdata.MO@meta.data[,c("eq_subclass")])
colnames(MO.nt)<-"eq_subclass"
MO.nt <- MO.nt[duplicated(MO.nt$eq_subclass)==F, ]
MO.nt <- data.frame(subclass=MO.nt, Glut=rep(0, length(MO.nt)), Gaba=rep(0, length(MO.nt)), Dopa=rep(0, length(MO.nt)), Chol=rep(0, length(MO.nt)), Sert=rep(0, length(MO.nt)), Nora=rep(0, length(MO.nt)), Gly=rep(0, length(MO.nt)), Hist=rep(0, length(MO.nt)), Neurotransmitter=rep(0, length(MO.nt)))

#find an arbituary threshold for neurotransmitter markers.slow step

#Dopaminergic
#Dopaminergic (Dopa): (Slc6a3 (Dat) or Slc18a2) and (Th and Ddc).
MO.NT.plot.Dopa <- DotPlot(scdata.MO,c("Slc6a3","Slc18a2", "Ddc","TH"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc6a3.exp <- NT.marker.threshold("Slc6a3", scdata.MO, "eq_subclass")
ddc.exp <- NT.marker.threshold("Ddc", scdata.MO, "eq_subclass")
th.exp <- NT.marker.threshold("TH", scdata.MO, "eq_subclass")
slc18a2.exp <- NT.marker.threshold("Slc18a2", scdata.MO, "eq_subclass")

dopa.all.exp<-data.frame(celltype=slc6a3.exp$celltype, slc6a3=slc6a3.exp$mean_lnCPM, slc18a2=slc18a2.exp$mean_lnCPM, ddc=ddc.exp$mean_lnCPM, th=th.exp$mean_lnCPM)
# Save the scatter plot in a variable
p1 <- ggplot(dopa.all.exp, aes(x = slc18a2, y = th)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p1, type = "histogram")
p1 <- ggplot(dopa.all.exp, aes(x = slc6a3, y = th)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p1, type = "histogram")
p1 <- ggplot(dopa.all.exp, aes(x = ddc, y = th)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p1, type = "histogram")

slc6a3 <- slc6a3.exp[slc6a3.exp[,2]>0.2,1]
slc18a2 <- slc18a2.exp[slc18a2.exp[,2]>3,1]
th <- th.exp[th.exp[,2]>1,1]
ddc <- ddc.exp[ddc.exp[,2]>3, 1]
dopa <- intersect(union(slc6a3,slc18a2), intersect(ddc, th))
MO.nt$Dopa[MO.nt$subclass %in% dopa]<-1

#Gabaergic
#GABAergic (GABA): (Slc32a1 (Vgat) or Slc18a2 (Vmat2)) and (Gad1, Gad2 or Aldh1a1).
MO.NT.plot.Gaba <- DotPlot(scdata.MO,c("Slc32a1","Slc18a2","Gad1", "Gad2","Slc17a6","Aldh1a1"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
slc32a1.exp <- NT.marker.threshold("Slc32a1", scdata.MO, "eq_subclass")
gad1.exp <- NT.marker.threshold("Gad1", scdata.MO, "eq_subclass")
gad2.exp <- NT.marker.threshold("Gad2", scdata.MO, "eq_subclass")
aldh1a1.exp <- NT.marker.threshold("Aldh1a1", scdata.MO, "eq_subclass")
slc17a6.exp <- NT.marker.threshold("Slc17a6", scdata.MO, "eq_subclass")
gaba.all.exp <- data.frame(
  celltype = slc32a1.exp$celltype, 
  slc32a1  = slc32a1.exp$mean_lnCPM, 
  gad1     = gad1.exp$mean_lnCPM, 
  gad2     = gad2.exp$mean_lnCPM, 
  aldh1a1  = aldh1a1.exp$mean_lnCPM,  
  slc17a6  = slc17a6.exp$mean_lnCPM
)
# Save the scatter plot in a variable
p1 <- ggplot(gaba.all.exp, aes(x = slc17a6, y = slc32a1)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p1, type = "histogram")

slc32a1 <- slc32a1.exp[slc32a1.exp[,2]>0.15,1]
p2 <- ggplot(gaba.all.exp, aes(x = slc17a6, y = gad1)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p2, type = "histogram")

gad1 <- gad1.exp[gad1.exp[,2]>1.5,1]

p3 <- ggplot(gaba.all.exp, aes(x = slc17a6, y = gad2)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p3, type = "histogram")
gad2 <- gad2.exp[gad2.exp[,2]>2,1]


aldh1a1 <- aldh1a1.exp[aldh1a1.exp[,2]>1,1]

gaba <- intersect(union(slc18a2, slc32a1), union(gad1,union(gad2, aldh1a1)))
gaba[order(gaba)]
MO.nt$Gaba[MO.nt$subclass %in% gaba]<-1


#Glutamatergic
#Glutamatergic (Glut): Slc17a6 (also known as Vglut2), Slc17a7 (Vglut1) or Slc17a8 (Vglut3).
MO.nt.plot.Glut <- DotPlot(scdata.MO,c("Slc17a6","Slc17a7","Slc17a8", "Slc32a1"), scale.max = 20,cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc17a6.exp <- NT.marker.threshold("Slc17a6", scdata.MO, "eq_subclass")
slc17a7.exp <- NT.marker.threshold("Slc17a7", scdata.MO, "eq_subclass")
slc17a8.exp <- NT.marker.threshold("Slc17a8", scdata.MO, "eq_subclass")

gaba.all.exp<-data.frame(celltype=slc32a1.exp$celltype, slc32a1=slc32a1.exp$mean_lnCPM, gad1=gad1.exp$mean_lnCPM, gad2=gad2.exp$mean_lnCPM, slc17a6=slc17a6.exp$mean_lnCPM,slc17a7=slc17a7.exp$mean_lnCPM, slc17a8=slc17a8.exp$mean_lnCPM)
p1 <- ggplot(gaba.all.exp, aes(x = slc17a8, y = slc32a1)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p1, type = "histogram")

slc17a6 <- slc17a6.exp[slc17a6.exp[,2]>0.5,1]
slc17a7 <- slc17a7.exp[slc17a7.exp[,2]>1,1]
slc17a8 <- slc17a8.exp[slc17a8.exp[,2]>1,1]
glut <- union(slc17a6, union(slc17a7, slc17a8))
glut[order(glut)]
MO.nt$Glut[MO.nt$subclass %in% glut]<-1


#Serotoninergic
#Serotonergic (Sero): (Slc6a4 (Sert) or Slc18a2) and (Tph2 and Ddc).
MO.NT.plot.Sert <- DotPlot(scdata.MO,c("Slc6a4","Slc18a2","Tph2", "Ddc"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
tph2.exp <- NT.marker.threshold("Tph2", scdata.MO, "eq_subclass")
slc6a4.exp <- NT.marker.threshold("Slc6a4", scdata.MO, "eq_subclass")
slc6a4 <- slc6a4.exp[slc6a4.exp[,2]>3,1]
tph2 <- tph2.exp[tph2.exp[,2]>3,1]
sert <- intersect(union(slc6a4, slc18a2), intersect(tph2, ddc))
MO.nt$Sert[MO.nt$subclass %in% sert]<-1

#Glycinergic
#Glycinergic (Glyc): Slc6a5.
MO.NT.plot.Gly <- DotPlot(scdata.MO,c("Slc6a5"),cols = c("lightgrey","darkred"), scale.max = 20)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
slc6a5.exp <- NT.marker.threshold("Slc6a5", scdata.MO, "eq_subclass")
gly <- slc6a5.exp[slc6a5.exp[,2]>0.5,1]
MO.nt$Gly[MO.nt$subclass %in% gly]<-1

#Noradrenergic
#Noradrenergic (Nora): (Slc6a2 (Net) or Slc18a2) and Dbh.Slc6a2 is not expressed.
MO.NT.plot.Nora <- DotPlot(scdata.MO,c("Slc6a2","Slc18a2","Dbh"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
dbh.exp <- NT.marker.threshold("Dbh", scdata.MO, "eq_subclass")
dbh <- dbh.exp[dbh.exp[,2]>1,1]
nora <- intersect(slc18a2, dbh)
MO.nt$Nora[MO.nt$subclass %in% nora]<-1


#Cholinergic
#Cholinergic (Chol): Slc18a3 (Vacht) and Chat. Slc18a3 is not in the annotation.
MO.NT.plot.Chol <- DotPlot(scdata.MO,c("Chat"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
chat.exp <- NT.marker.threshold("Chat", scdata.MO, "eq_subclass")
chat <- chat.exp[chat.exp[,2]>3,1]
chol <- chat
MO.nt$Chol[MO.nt$subclass %in% chol]<-1


#histaminergic
#Histaminergic (Hist): Slc18a2 and Hdc.
MO.NT.plot.Hist <- DotPlot(scdata.MO,c("Slc18a2","Hdc"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
hdc.exp <- NT.marker.threshold("Hdc", scdata.MO, "eq_subclass")
hdc <- hdc.exp[hdc.exp[,2]>3,1]
hist <- intersect(slc18a2, hdc)
MO.nt$Hist[MO.nt$subclass %in% hist]<-1

NT <- c("Glut","Gaba", "Dopa", "Chol", "Sert", "Nora", "Gly", "Hist")
NT.1 <- t(replicate(nrow(MO.nt), NT))
NT.2 <- t(apply(MO.nt[,2:9], 1, as.logical))
for (i in 1:nrow(MO.nt)) {
  MO.nt$Neurotransmitter[i] <- paste(NT.1[i,][NT.2[i,]], collapse="-")
}

MO.nt$Neurotransmitter[MO.nt$Neurotransmitter==""]<-"unclear"
#future(saveRDS(scdata.MO, file="scdata.MO.rds"))

#save data
write.csv(MO.nt, file="./MO.nt.type.csv")
MO.nt<-read.csv(file="./MO.nt.type.csv")
#add to metadata
scdata.MO@meta.data$neurotransmitter <- MO.nt$Neurotransmitter[match(scdata.MO@meta.data$eq_subclass, MO.nt$subclass)]
color <- colorRampPalette(brewer.pal(9, "Set1")[1:8])(13)
color <- c("grey",color)
DimPlot(scdata.MO, group.by = "neurotransmitter", cols = color, reduction = "tsne")+coord_fixed()
#future(saveRDS(scdata.MO, file="scdata.MO.rds"))
qs_save(scdata.MO, file = "scdata.MO.qs2", nthreads = 8)
# Read it back in using 4 CPU cores
scdata.MO <- qs_read("scdata.MO.qs2", nthreads = 8)
View(MO.nt)
```


## Annotate Green anole NT type

```{r}
Idents(scdata.AC)<-"eq_subclass"
AC.nt <- data.frame(scdata.AC@meta.data[,c("eq_subclass")])
colnames(AC.nt)<-"eq_subclass"
AC.nt <- AC.nt[duplicated(AC.nt$eq_subclass)==F, ]
AC.nt <- data.frame(subclass=AC.nt, Glut=rep(0, length(AC.nt)), Gaba=rep(0, length(AC.nt)), Dopa=rep(0, length(AC.nt)), Chol=rep(0, length(AC.nt)), Sert=rep(0, length(AC.nt)), Nora=rep(0, length(AC.nt)), Gly=rep(0, length(AC.nt)), Hist=rep(0, length(AC.nt)), Neurotransmitter=rep(0, length(AC.nt)))

#find an arbituary threshold for neurotransmitter markers.

#Dopaminergic
#Dopaminergic (Dopa): (Slc6a3 (Dat) or Slc18a2) and (Th and Ddc). Dat not in the annotation
AC.NT.plot.Dopa <- DotPlot(scdata.AC,c("slc18a2", "ddc","th"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
ddc.exp <- NT.marker.threshold("ddc", scdata.AC, "eq_subclass")
th.exp <- NT.marker.threshold("th", scdata.AC, "eq_subclass")
slc18a2.exp <- NT.marker.threshold("slc18a2", scdata.AC, "eq_subclass")
slc18a2 <- slc18a2.exp[slc18a2.exp[,2]>2,1]
th <- th.exp[th.exp[,2]>3,1]
ddc <- ddc.exp[ddc.exp[,2]>3,1]
dopa <- intersect(slc18a2, intersect(ddc, th))
AC.nt$Dopa[AC.nt$subclass %in% dopa]<-1

#Gabaergic
#GABAergic (GABA): (Slc32a1 (Vgat) or Slc18a2 (Vmat2)) and (Gad1, Gad2 or Aldh1a1). 
#Glutamatergic
#Glutamatergic (Glut): Slc17a6 (also known as Vglut2), Slc17a7 (Vglut1) or Slc17a8 (Vglut3).


AC.NT.plot.Gaba <- DotPlot(scdata.AC, c("slc32a1","slc18a2", "gad1","gad2","slc17a6","slc17a7","slc17a8","aldh1a1"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc32a1.exp <- NT.marker.threshold("slc32a1", scdata.AC, "eq_subclass")
gad2.exp <- NT.marker.threshold("gad2", scdata.AC, "eq_subclass")
gad1.exp <- NT.marker.threshold("gad1", scdata.AC, "eq_subclass")
aldh1a1.exp <- NT.marker.threshold("aldh1a1", scdata.AC, "eq_subclass")
slc17a6.exp <- NT.marker.threshold("slc17a6", scdata.AC, "eq_subclass")
slc17a7.exp <- NT.marker.threshold("slc17a7", scdata.AC, "eq_subclass")
slc17a8.exp <- NT.marker.threshold("slc17a8", scdata.AC, "eq_subclass")


gaba.all.exp<-data.frame(celltype=slc32a1.exp$celltype, slc32a1=slc32a1.exp$mean_lnCPM, gad1=gad1.exp$mean_lnCPM, gad2=gad2.exp$mean_lnCPM, slc17a6=slc17a6.exp$mean_lnCPM,slc17a7=slc17a7.exp$mean_lnCPM, slc17a8=slc17a8.exp$mean_lnCPM)


p1 <- ggplot(gaba.all.exp, aes(x = slc32a1, y = slc17a6)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p1, type = "histogram")

slc32a1 <- slc32a1.exp[slc32a1.exp[,2]>0.3,1]
gad1 <- gad1.exp[gad1.exp[,2]>1.5,1]
gad2 <- gad2.exp[gad2.exp[,2]>2,1]
aldh1a1 <- aldh1a1.exp[aldh1a1.exp[,2]>1,1]
gaba <- intersect(union(slc18a2, slc32a1), union(gad1,union(gad2, aldh1a1)))
AC.nt$Gaba[AC.nt$subclass %in% gaba]<-1

slc17a6 <- slc17a6.exp[slc17a6.exp[,2]>1,1]
slc17a7 <- slc17a7.exp[slc17a7.exp[,2]>0.5,1]
slc17a8 <- slc17a8.exp[slc17a8.exp[,2]>0.5,1]
glut <- union(slc17a6, union(slc17a7, slc17a8))
AC.nt$Glut[AC.nt$subclass %in% glut]<-1


#Serotoninergic
#Serotonergic (Sero): (Slc6a4 (Sert) or Slc18a2) and (Tph2 and Ddc).
AC.NT.plot.Sert <- DotPlot(scdata.AC,c("slc6a4","slc18a2","tph2", "ddc","tph1"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
tph2.exp <- NT.marker.threshold("tph2", scdata.AC, "eq_subclass")
tph1.exp <- NT.marker.threshold("tph1", scdata.AC, "eq_subclass")
slc6a4.exp <- NT.marker.threshold("slc6a4", scdata.AC, "eq_subclass")
slc6a4 <- slc6a4.exp[slc6a4.exp[,2]>3,1]
tph2 <- tph2.exp[tph2.exp[,2]>3,1]
tph1 <- tph1.exp[tph1.exp[,2]>0.3,1]
sert <- intersect(union(slc6a4, slc18a2), intersect(union(tph2,tph1), ddc))
AC.nt$Sert[AC.nt$subclass %in% sert]<-1

#Glycinergic
#Glycinergic (Glyc): Slc6a5.
AC.NT.plot.Gly <- DotPlot(scdata.AC,c("slc6a5"),cols = c("lightgrey","darkred"),scale.max = 20)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc6a5.exp <- NT.marker.threshold("slc6a5", scdata.AC, "eq_subclass")
gly <- slc6a5.exp[slc6a5.exp[,2]>0.3,1]
AC.nt$Gly[AC.nt$subclass %in% gly]<-1

#Noradrenergic
#Noradrenergic (Nora): (Slc6a2 (Net) or Slc18a2) and Dbh Slc6a2 not in the genome
AC.NT.plot.Nora <- DotPlot(scdata.AC,c("slc6a2","slc18a2","dbh"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc6a2.exp <- NT.marker.threshold("slc6a2", scdata.AC, "eq_subclass")
dbh.exp <- NT.marker.threshold("dbh", scdata.AC, "eq_subclass")
dbh <- dbh.exp[dbh.exp[,2]>1,1]
slc6a2 <- slc6a2.exp[slc6a2.exp[,2]>1,1]
nora <- intersect(slc18a2, dbh)

#Cholinergic
#Cholinergic (Chol): Slc18a3 (Vacht) and Chat. Slc18a3 is not in the annotation.
AC.NT.plot.Chol <- DotPlot(scdata.AC,c("slc18a3","chat"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
chat.exp <- NT.marker.threshold("chat", scdata.AC, "eq_subclass")
slc18a3.exp <- NT.marker.threshold("slc18a3", scdata.AC, "eq_subclass")
chat <- chat.exp[chat.exp[,2]>3,1]
slc18a3 <- slc18a3.exp[slc18a3.exp[,2]>3,1]
chol <- intersect(chat,slc18a3)
AC.nt$Chol[AC.nt$subclass %in% chol]<-1


#histaminergic
#Histaminergic (Hist): Slc18a2 and Hdc.
AC.NT.plot.Hist <- DotPlot(scdata.AC,c("slc18a2","hdc"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
hdc.exp <- NT.marker.threshold("hdc", scdata.AC, "eq_subclass")
hdc <- hdc.exp[hdc.exp[,2]>0.5,1]
hist <- intersect(slc18a2, hdc)
AC.nt$Hist[AC.nt$subclass %in% hist]<-1

NT <- c("Glut","Gaba", "Dopa", "Chol", "Sert", "Nora", "Gly", "Hist")
NT.1 <- t(replicate(nrow(AC.nt), NT))
NT.2 <- t(apply(AC.nt[,2:9], 1, as.logical))
for (i in 1:nrow(AC.nt)) {
  AC.nt$Neurotransmitter[i] <- paste(NT.1[i,][NT.2[i,]], collapse="-")
}

AC.nt$Neurotransmitter[AC.nt$Neurotransmitter==""]<-"unclear"


#save data
write.csv(AC.nt, file="./AC.nt.type.csv")
AC.nt<-read.csv(file="./AC.nt.type.csv")
#add to metadata
scdata.AC@meta.data$neurotransmitter <- AC.nt$Neurotransmitter[match(scdata.AC@meta.data$eq_subclass, AC.nt$subclass)]
color=c("#FF3358","#73E785","#820e57","#2B93DF","#fcf04b","#0a9964","#820e57","#ff7621","#820e57","#4F2E91FF","#99999960")
DimPlot(scdata.AC, group.by = "neurotransmitter", cols = color, reduction = "tsne")+coord_fixed()
aggregate(rep(1, nrow(scdata.AC@meta.data)), by = list(x = scdata.AC@meta.data$neurotransmitter, y = scdata.AC@meta.data$eq_subclass), sum)
#future(saveRDS(scdata.AC, file="scdata.AC.rds"))
qs_save(scdata.AC, file = "scdata.AC.qs2", nthreads = 8)
# Read it back in using 4 CPU cores
scdata.AC <- qs_read("scdata.AC.qs2", nthreads = 8)
View(AC.nt)
```


## Annotate Zebrafish NT type
```{r}
Idents(scdata.DR)<-"eq_subclass"
DR.nt <- data.frame(scdata.DR@meta.data[,c("eq_subclass")])
colnames(DR.nt)<-"eq_subclass"
DR.nt <- DR.nt[duplicated(DR.nt$eq_subclass)==F, ]
DR.nt <- data.frame(subclass=DR.nt, Glut=rep(0, length(DR.nt)), Gaba=rep(0, length(DR.nt)), Dopa=rep(0, length(DR.nt)), Chol=rep(0, length(DR.nt)), Sert=rep(0, length(DR.nt)), Nora=rep(0, length(DR.nt)), Gly=rep(0, length(DR.nt)), Hist=rep(0, length(DR.nt)), Neurotransmitter=rep(0, length(DR.nt)))

#find an arbituary threshold for neurotransmitter markers.

#Dopaminergic
#Dopaminergic (Dopa): (Slc6a3 (Dat ENSDARG00000004219.10) or Slc18a2 ENSDARG00000015110.10) and (Th ENSDARG00000030621.8 and Ddc ENSDARG00000016494.9). Dat not in the annotation
DR.NT.plot.Dopa <- DotPlot(scdata.DR,c("slc6a3","ddc", "th","th-1","slc18a2","slc18a2-1"),cols = c("lightgrey","darkred"), scale.max = 20)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 10), axis.text.y = element_text(size = 10))
slc6a3.exp <- NT.marker.threshold("slc6a3", scdata.DR, "eq_subclass")
ddc.exp <- NT.marker.threshold("ddc", scdata.DR, "eq_subclass")
#th.exp <- NT.marker.threshold("th", scdata.DR, "eq_subclass")
th_1.exp <- NT.marker.threshold("th-1", scdata.DR, "eq_subclass")
slc18a2.exp <- NT.marker.threshold("slc18a2", scdata.DR, "eq_subclass")
slc18a2_1.exp <- NT.marker.threshold("slc18a2-1", scdata.DR, "eq_subclass")
slc6a3 <- slc6a3.exp[slc6a3.exp[,2]>1,1]
slc18a2 <- slc18a2.exp[slc18a2.exp[,2]>0.1,1]
slc18a2_1 <- slc18a2_1.exp[slc18a2_1.exp[,2]>1,1]
#th <- th.exp[th.exp[,2]>1,1]
th_1 <- th_1.exp[th_1.exp[,2]>1,1]
ddc <- ddc.exp[ddc.exp[,2]>2,1]
dopa <- intersect(union(union(slc6a3,slc18a2),slc18a2_1), intersect(ddc, th_1))
DR.nt$Dopa[DR.nt$subclass %in% dopa]<-1




#Gabaergic
#GABAergic (GABA): (Slc32a1 (Vgat) or Slc18a2 (Vmat2)) and (Gad1, Gad2 or Aldh1a1). 
#Glutamatergic
#Glutamatergic (Glut): Slc17a6 (also known as Vglut2), Slc17a7 (Vglut1) or Slc17a8 (Vglut3).


DR.NT.plot.Gaba <- DotPlot(scdata.DR, c("slc32a1","slc18a2", "gad1","gad2","slc17a6","slc17a7","slc17a8","aldh1a1"),cols = c("lightgrey","darkred"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc32a1.exp <- NT.marker.threshold("slc32a1", scdata.DR, "eq_subclass")
gad2.exp <- NT.marker.threshold("gad2", scdata.DR, "eq_subclass")
gad1a.exp <- NT.marker.threshold("gad1a", scdata.DR, "eq_subclass")
gad1b.exp <- NT.marker.threshold("gad1b", scdata.DR, "eq_subclass")
#gad3.exp <- NT.marker.threshold("gad3", scdata.DR, "eq_subclass")
slc17a6a.exp <- NT.marker.threshold("slc17a6a", scdata.DR, "eq_subclass")
#aldh1a1.exp <- NT.marker.threshold("aldh1a1", scdata.DR, "eq_subclass") Teleost does not have this gene
slc17a6a.exp <- NT.marker.threshold("slc17a6a", scdata.DR, "eq_subclass")
slc17a6b.exp <- NT.marker.threshold("slc17a6b", scdata.DR, "eq_subclass")
#slc17a7a.exp <- NT.marker.threshold("slc17a7a", scdata.DR, "eq_subclass")
#slc17a7a_1.exp <- NT.marker.threshold("slc17a7a-1", scdata.DR, "eq_subclass")
slc17a7b.exp <- NT.marker.threshold("slc17a7b", scdata.DR, "eq_subclass")
slc17a8.exp <- NT.marker.threshold("slc17a8", scdata.DR, "eq_subclass")
#slc17a8_1.exp <- NT.marker.threshold("slc17a8-1", scdata.DR, "eq_subclass")


gaba.all.exp<-data.frame(celltype=slc32a1.exp$celltype, slc32a1=slc32a1.exp$mean_lnCPM, gad1a=gad1a.exp$mean_lnCPM,  gad1b=gad1b.exp$mean_lnCPM, gad2=gad2.exp$mean_lnCPM, slc17a6a=slc17a6a.exp$mean_lnCPM, slc17a6b=slc17a6b.exp$mean_lnCPM,  slc17a7b=slc17a7b.exp$mean_lnCPM, slc17a8=slc17a8.exp$mean_lnCPM)


p1 <- ggplot(gaba.all.exp, aes(x = slc32a1, y = slc17a6a)) +
  geom_point()
# Plot the scatter plot with marginal histograms
ggMarginal(p1, type = "histogram")


slc32a1 <- slc32a1.exp[slc32a1.exp[,2]>2,1]
gad1a <- gad1a.exp[gad1a.exp[,2]>1,1]
gad2 <- gad2.exp[gad2.exp[,2]>2,1]
gad1b <- gad1b.exp[gad1b.exp[,2]>1,1]
#gad3 <- gad3.exp[gad3.exp[,2]>1,1]
gaba <- intersect(union(slc18a2, slc32a1), union(union(gad1a, gad2),gad1b))
DR.nt$Gaba[DR.nt$subclass %in% gaba]<-1


slc17a6a <- slc17a6a.exp[slc17a6a.exp[,2]>0.1,1]
slc17a6b <- slc17a6b.exp[slc17a6b.exp[,2]>0.2,1]
#slc17a7a <- slc17a7a.exp[slc17a7a.exp[,2]>1,1]
#slc17a7a_1 <- slc17a7a_1.exp[slc17a7a_1.exp[,2]>0.1,1]
slc17a7b <- slc17a7b.exp[slc17a7b.exp[,2]>0.15,1]
slc17a8 <- slc17a8.exp[slc17a8.exp[,2]>0.5,1]
#slc17a8_1 <- slc17a8_1.exp[slc17a8_1.exp[,2]>0.5,1]
glut <- union(slc17a6b, union(slc17a6a, union(slc17a7b, slc17a8)))
DR.nt$Glut[DR.nt$subclass %in% glut]<-1




#Serotoninergic
#Serotonergic (Sero): (Slc6a4 (Sert) a ENSDARG00000061165.6 b ENSDARG00000057017.6 or Slc18a2 ENSDARG00000015110.10) and (Tph2 ENSDARG00000057239.7 and Ddc ENSDARG00000016494.9). tph1b ENSDARG00000036082.5 tph1a ENSDARG00000029432.8 
# sert b is more correlated with tph2 and ddc. 
# tph1a is more correlated with ddc
# the serotoninergic neurons includes tph1 positive cells in fish
DR.NT.plot.Sert <- DotPlot(scdata.DR,c("slc6a4a","slc6a4b", "slc18a2","slc18a2-1","tph1a","tph1b","tph1a-1","tph1b-1","tph2","tph2-1","ddc"),cols = c("lightgrey","darkred"), scale.max = 20)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 10), axis.text.y = element_text(size = 10))
#slc6a4a.exp <- NT.marker.threshold("slc6a4a", scdata.DR, "eq_subclass")
slc6a4b.exp <- NT.marker.threshold("slc6a4b", scdata.DR, "eq_subclass")
tph2.exp <- NT.marker.threshold("tph2", scdata.DR, "eq_subclass")
tph2_1.exp <- NT.marker.threshold("tph2-1", scdata.DR, "eq_subclass")
tph1a.exp <- NT.marker.threshold("tph1a", scdata.DR, "eq_subclass")
tph1a_1.exp <- NT.marker.threshold("tph1a-1", scdata.DR, "eq_subclass")
tph1b.exp <- NT.marker.threshold("tph1b", scdata.DR, "eq_subclass")
tph1b_1.exp <- NT.marker.threshold("tph1b-1", scdata.DR, "eq_subclass")
ddc.exp <- NT.marker.threshold("ddc", scdata.DR, "eq_subclass")
#slc6a4a <- slc6a4a.exp[slc6a4a.exp[,2]>0.1,1]
slc6a4b <- slc6a4b.exp[slc6a4b.exp[,2]>1,1]
tph2 <- tph2.exp[tph2.exp[,2]>1,1]
tph2_1 <- tph2_1.exp[tph2_1.exp[,2]>1,1]
tph1a <- tph1a.exp[tph1a.exp[,2]>1,1]
tph1a_1 <- tph1a_1.exp[tph1a_1.exp[,2]>0.5,1]
tph1b <- tph1b.exp[tph1b.exp[,2]>1,1]
tph1b_1 <- tph1b_1.exp[tph1b_1.exp[,2]>1,1]
sert <- intersect(union(slc6a4b, slc18a2), intersect(union(tph2_1, union(tph1b_1,union(tph1a_1,union(tph1b,union(tph1a,tph2))))), ddc))
DR.nt$Sert[DR.nt$subclass %in% sert]<-1

#Glycinergic
#Glycinergic (Glyc): Slc6a5 ENSDARG00000067964 , not found in data. but the glia glycine transporter is found:ENSDARG00000018534.13
DR.NT.plot.Gly <- DotPlot(scdata.DR,c("slc6a5"),cols = c("lightgrey","darkred"),scale.max = 20)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))

#don't exist

#Noradrenergic
#Noradrenergic (Nora): (Slc6a2 (Net) ENSDARG00000016141.13 or Slc18a2 ENSDARG00000015110.10) and Dbh ENSDARG00000069446.6
DR.NT.plot.Nora <- DotPlot(scdata.DR,c("slc6a2","slc18a2","slc18a2-1","dbh","dbh-1"),cols = c("lightgrey","darkred"),scale.max = 20)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
slc6a2.exp <- NT.marker.threshold("slc6a2", scdata.DR, "eq_subclass")
dbh.exp <- NT.marker.threshold("dbh", scdata.DR, "eq_subclass")
#dbh_1.exp <- NT.marker.threshold("dbh-1", scdata.DR, "eq_subclass")
slc6a2 <- slc6a2.exp[slc6a2.exp[,2]>3,1]
dbh <- dbh.exp[dbh.exp[,2]>1,1]
#dbh_1 <- dbh_1.exp[dbh_1.exp[,2]>1,1]
nora <- intersect(slc6a2, dbh)
DR.nt$Nora[DR.nt$subclass %in% nora]<-1
#don't exist

#Cholinergic
#Cholinergic (Chol): Slc18a3 (Vacht) a ENSDARG00000006356.6 b ENSDARG00000107527.2 and Chat ENSDARG00000015854.11. 
DR.NT.plot.Chol <- DotPlot(scdata.DR,c("slc18a3a", "slc18a3b","chata"),cols = c("lightgrey","darkred"), scale.max = 20)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5), axis.text.y = element_text(size = 5))
chat.exp <- NT.marker.threshold("chata", scdata.DR, "eq_subclass")
slc18a3a.exp <- NT.marker.threshold("slc18a3a", scdata.DR, "eq_subclass")
slc18a3b.exp <- NT.marker.threshold("slc18a3b", scdata.DR, "eq_subclass")
chat <- chat.exp[chat.exp[,2]>1,1]
slc18a3a <- slc18a3a.exp[slc18a3a.exp[,2]>1,1]
slc18a3b <- slc18a3b.exp[slc18a3b.exp[,2]>1,1]
chol <- intersect(chat,union(slc18a3a,slc18a3b))
DR.nt$Chol[DR.nt$subclass %in% chol]<-1


#histaminergic
#Histaminergic (Hist): Slc18a2 ENSDARG00000015110.10 and Hdc ENSDARG00000075454.6 .
DR.NT.plot.Hist <- DotPlot(scdata.DR,c("slc18a2","slc18a2-1","hdc"),cols = c("lightgrey","darkred"), scale.max = 20, assay = "RNA")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 10), axis.text.y = element_text(size = 10))
hdc.exp <- NT.marker.threshold("hdc", scdata.DR, "eq_subclass")
hdc <- hdc.exp[hdc.exp[,2]>1,1]
hist <- intersect(union(slc18a2_1,slc18a2), hdc)
DR.nt$Hist[DR.nt$subclass %in% hist]<-1

NT <- c("Glut","Gaba", "Dopa", "Chol", "Sert", "Nora", "Gly", "Hist")
NT.1 <- t(replicate(nrow(DR.nt), NT))
NT.2 <- t(apply(DR.nt[,2:9], 1, as.logical))
for (i in 1:nrow(DR.nt)) {
  DR.nt$Neurotransmitter[i] <- paste(NT.1[i,][NT.2[i,]], collapse="-")
}

DR.nt$Neurotransmitter[DR.nt$Neurotransmitter==""]<-"unclear"


#save data
write.csv(DR.nt, file="./DR.nt.type.csv")
DR.nt<-read.csv(file="./DR.nt.type.csv")
#add to metadata
scdata.DR@meta.data$neurotransmitter <- DR.nt$Neurotransmitter[match(scdata.DR@meta.data$eq_subclass, DR.nt$subclass)]
color <- colorRampPalette(brewer.pal(9, "Set1")[1:8])(9)
color <- c(color,"grey")
color=c("#FF3358","#fcf04b","#4F2E91FF","#ff7621","#2B93DF","#73E785","#0a9964",
        "#99999960")
DimPlot(scdata.DR, group.by = "neurotransmitter", cols = color, reduction = "tsne")+coord_fixed()
#future(saveRDS(scdata.DR, file="scdata.DR.rds"))
qs_save(scdata.DR, file = "scdata.DR.qs2", nthreads = 8)
# Read it back in using 4 CPU cores
#scdata.DR <- qs_read("scdata.DR.qs2", nthreads = 8)
View(DR.nt)
```














## Examine all possible combination of types between species difference
```{r}

scdata.MO <- qs_read("scdata.MO.qs2", nthreads = 8)
scdata.CJ <- qs_read("scdata.CJ.qs2", nthreads = 8)
scdata.AC <- qs_read("scdata.AC.qs2", nthreads = 8)
scdata.XT <- qs_read("scdata.XT.qs2", nthreads = 8)
scdata.DR <- qs_read("scdata.DR.qs2", nthreads = 8)

scdata.AI<-readRDS(file="scdata.AI.HY.rds")
Idents(scdata.AI)<-"class"
scdata.AI.HYsubclass <- subset(scdata.AI, idents =  c("11 CNU-HYa GABA","12 HY GABA","13 CNU-HYa Glut","14 HY Glut"))
unique(scdata.AI.HYsubclass@meta.data$neurotransmitter)
unique(scdata.MO@meta.data$neurotransmitter)
unique(scdata.CJ@meta.data$neurotransmitter)
unique(scdata.AC@meta.data$neurotransmitter)
unique(scdata.XT@meta.data$Neurotransmitter)
unique(scdata.DR@meta.data$neurotransmitter)


```



## Table plot of NT changes

```{r}
metadata.AI.HYsubclass <- scdata.AI.HYsubclass@meta.data
metadata.AI.HYsubclass <- metadata.AI.HYsubclass[metadata.AI.HYsubclass$neurotransmitter != "",]
AI.NT <- aggregate(rep(1, nrow(metadata.AI.HYsubclass)), by = list(x = metadata.AI.HYsubclass$neurotransmitter, y = metadata.AI.HYsubclass$subclass), sum)
# Each subclass can contain more than one neurotransmitter types. use the union for following comparisons.
# "HY" means cells collected from HY dissection. It includes 192 subclasses. 80 belongs to the 4 major HY classes.
colnames(AI.NT) <- c("neurotransmitter", "subclass", "#cells")
AI.NT <- AI.NT[duplicated(AI.NT$subclass)==F,]
AI.NT[AI.NT$subclass=="058 PAL-STR Gaba-Chol",1]<-"GABA-Chol"
AI.NT[AI.NT$subclass=="081 ACB-BST-FS D1 Gaba",1]<-"Glut-GABA"
AI.NT[AI.NT$subclass=="085 SI-MPO-LPO Lhx8 Gaba",1]<-"Glut-GABA"
AI.NT[AI.NT$subclass=="089 PVR Six3 Sox3 Gaba",1]<-"GABA-Dopa"
AI.NT[AI.NT$subclass=="092 TMv-PMv Tbx3 Hist-Gaba",1]<-"GABA-Hist"
AI.NT[AI.NT$subclass=="098 AHN-SBPV-PVHd Pdrm12 Gaba",1]<-"GABA-Dopa"
AI.NT[AI.NT$subclass=="099 SBPV-PVa Six6 Satb2 Gaba",1]<-"GABA-Dopa"
AI.NT[AI.NT$subclass=="101 ZI Pax6 Gaba",1]<-"GABA-Dopa"
AI.NT[AI.NT$subclass=="102 DMH-LHA Gsx1 Gaba",1]<-"GABA-Chol"
AI.NT[AI.NT$subclass=="106 PVpo-VMPO-MPN Hmx2 Gaba",1]<-"GABA-Dopa"
AI.NT[AI.NT$subclass=="107 DMH Hmx2 Gaba",1]<-"Glut-GABA-Dopa"
AI.NT[AI.NT$subclass=="115 MS-SF Bsx Glut",1]<-"Glut-GABA"
AI.NT[AI.NT$subclass=="119 SI-MA-LPO-LHA Skor1 Glut",1]<-"Glut-GABA"
AI.NT[AI.NT$subclass=="125 DMH Hmx2 Glut",1]<-"Glut-GABA"
AI.NT[AI.NT$subclass=="126 ARH-PVp Tbx3 Glut",1]<-"Glut-GABA"
AI.NT[AI.NT$subclass=="131 LHA-AHN-PVH Otp Trh Glut",1]<-"Glut-GABA"
AI.NT[AI.NT$subclass=="133 PVH-SO-PVa Otp Glut",1]<-"Glut-GABA"
AI.NT[AI.NT$subclass=="138 PH Pitx2 Glut",1]<-"Glut-Dopa"
AI.NT[AI.NT$subclass=="141 PH-SUM Foxa1 Glut",1]<-"Glut-GABA"
AI.NT[AI.NT$subclass=="145 MH Tac2 Glut",1]<-"Glut-Chol"
AI.NT[AI.NT$subclass=="159 IF-RL-CLI-PAG Foxa1 Glut",1]<-"Glut-Dopa"
AI.NT[AI.NT$subclass=="210 PRT Mecom Gaba",1]<-"GABA-Glyc"
AI.NT[AI.NT$subclass=="213 SCsg Gabrr2 Gaba",1]<-"Glut-GABA"
AI.NT[AI.NT$subclass=="215 SNc-VTA-RAmb Foxa1 Dopa",1]<-"Glut-GABA-Dopa"
AI.NT[AI.NT$subclass=="243 PGRN-PARN-MDRN Hoxb5 Glut",1]<-"Glut-GABA"
AI.NT[AI.NT$subclass=="285 MY Lhx1 Gly-Gaba",1]<-"GABA-Glyc"
#unify convention
AI.NT[AI.NT=="GABA"]<-"Gaba"
AI.NT[AI.NT=="GABA-Chol"]<-"Gaba-Chol"
AI.NT[AI.NT=="Glut-GABA"]<-"Glut-Gaba"
AI.NT[AI.NT=="GABA-Dopa"]<-"Gaba-Dopa"
AI.NT[AI.NT=="GABA-Hist"]<-"Gaba-Hist"
AI.NT[AI.NT=="Glut-GABA-Dopa"]<-"Glut-Gaba-Dopa"
AI.NT[AI.NT=="GABA-Glyc"]<-"Gaba-Gly"
colnames(AI.NT) <- c("MM.NT","subclass","MM_#cells")

#save data
write.csv(AI.NT, file="./AI.NT.type.csv")
AI.NT <- read.csv(file="./AI.NT.type.csv")

# extract HY subclasses.
#scdata.MO<-readRDS(file="scdata.MO.rds")
#scdata.CJ<-readRDS(file="scdata.CJ.rds")
#scdata.AC<-readRDS(file="scdata.AC.rds")
#scdata.XT<-readRDS(file="scdata.XT.rds")
#scdata.RI<-readRDS(file="scdata.RI.rds")
#scdata.DR<-readRDS(file="scdata.DR.rds")

MO.nt <- aggregate(rep(1, nrow(scdata.MO@meta.data)), by = list(x = scdata.MO@meta.data$neurotransmitter, 
                                                                y = scdata.MO@meta.data$eq_subclass), sum)
colnames(MO.nt) <- c("MO.NT", "subclass", "MO_#cells")
MO.nt.HY <- MO.nt[MO.nt$subclass %in% AI.NT$subclass,]

CJ.nt <- aggregate(rep(1, nrow(scdata.CJ@meta.data)), by = list(x = scdata.CJ@meta.data$neurotransmitter, 
                                                                y = scdata.CJ@meta.data$eq_subclass), sum)
colnames(CJ.nt) <- c("CJ.NT", "subclass", "CJ_#cells")
CJ.nt.HY <- CJ.nt[CJ.nt$subclass %in% AI.NT$subclass,]

AC.nt <- aggregate(rep(1, nrow(scdata.AC@meta.data)), by = list(x = scdata.AC@meta.data$neurotransmitter, 
                                                                y = scdata.AC@meta.data$eq_subclass), sum)
colnames(AC.nt) <- c("AC.NT", "subclass", "AC_#cells")
AC.nt.HY <- AC.nt[AC.nt$subclass %in% AI.NT$subclass,]

XT.nt <- aggregate(rep(1, nrow(scdata.XT@meta.data)), by = list(x = scdata.XT@meta.data$Neurotransmitter, 
                                                                y = scdata.XT@meta.data$eq_subclass), sum)
colnames(XT.nt) <- c("XT.NT", "subclass", "XT_#cells")
XT.nt.HY <- XT.nt[XT.nt$subclass %in% AI.NT$subclass,]

#RI.nt <- aggregate(rep(1, nrow(scdata.RI@meta.data)), by = list(x = scdata.RI@meta.data$neurotransmitter, 
#                                                                y = scdata.RI@meta.data$subclass_id_label_reduced_mapping_nounlabeled), sum)
#colnames(RI.nt) <- c("RI.NT", "subclass", "RI_#cells")
#RI.nt.HY <- RI.nt[RI.nt$subclass %in% AI.NT$subclass,]

DR.nt <- aggregate(rep(1, nrow(scdata.DR@meta.data)), by = list(x = scdata.DR@meta.data$neurotransmitter, 
                                                                y = scdata.DR@meta.data$eq_subclass), sum)
colnames(DR.nt) <- c("DR.NT", "subclass", "DR_#cells")
DR.nt.HY <- DR.nt[DR.nt$subclass %in% AI.NT$subclass,]


NT.all.mapped.HY<-merge(AI.NT, MO.nt.HY, by = "subclass", all = T)
NT.all.mapped.HY<-merge(NT.all.mapped.HY, CJ.nt.HY, by = "subclass", all = T)
NT.all.mapped.HY<-merge(NT.all.mapped.HY, AC.nt.HY, by = "subclass", all = T)
NT.all.mapped.HY<-merge(NT.all.mapped.HY, XT.nt.HY, by = "subclass", all = T)
NT.all.mapped.HY<-merge(NT.all.mapped.HY, DR.nt.HY, by = "subclass", all = T)

saveRDS(NT.all.mapped.HY, file="./NT.all.mapped.HT.rds")
NT.all.mapped.HY<-readRDS(file="./NT.all.mapped.HT.rds")

NT.switch.df <- NT.all.mapped.HY[,c("MM.NT","MO.NT","CJ.NT","AC.NT","XT.NT","DR.NT")]
colnames(NT.switch.df)<-c("Mousue","Prairie vole","Japanese quail", "Green anole","Tropical clawed frog", "zebrafish")
unique(unlist(NT.switch.df))

# numeric nt labels
# 1: Glut
# 2: Gaba
# 3,4,5,6: Dopa, including Dopa, Gaba-Dopa, GLut-Dopa, Glut-Gaba-Dopa
# 7,8,9, 10: Chol, including Chol, Gaba-Chol, GLut-Chol, Glut-Gaba-Chol,
# 11: Nora
# 12, 13, 14 : Gly
# 15: Hist
# 16: Glut-Gaba
# 0: unclear

NT.switch.df[NT.switch.df=="Glut"]<-1
NT.switch.df[NT.switch.df=="Gaba"]<-2
NT.switch.df[NT.switch.df=="Dopa"]<-3
NT.switch.df[NT.switch.df=="Gaba-Dopa"]<-3
NT.switch.df[NT.switch.df=="Glut-Dopa"]<-3
NT.switch.df[NT.switch.df=="Glut-Gaba-Dopa"]<-3
NT.switch.df[NT.switch.df=="Gaba-Chol"]<-4
NT.switch.df[NT.switch.df=="Nora"]<-5
NT.switch.df[NT.switch.df=="Gaba-Gly"]<-6
NT.switch.df[NT.switch.df=="Gaba-Hist"]<-7
NT.switch.df[NT.switch.df=="Glut-Gaba-Hist"]<-7
NT.switch.df[NT.switch.df=="Glut-Gaba-Sert-Hist"]<-7
NT.switch.df[NT.switch.df=="Glut-Gaba"]<-8
NT.switch.df[NT.switch.df=="unclear"]<-0
NT.switch.df[is.na(NT.switch.df)]<-9
NT.switch.df <- as.matrix(sapply(NT.switch.df, as.numeric))
rownames(NT.switch.df)<-NT.all.mapped.HY$subclass
color <- c("#00172D","#2B93DF","#FF3358","#fcf04b","#73E785","#03EDFF","#820e57","#ff7621","#0a9964","#999999")
switch.plot<-pheatmap(NT.switch.df,
         cluster_cols = FALSE,
         cluster_rows = FALSE,
         color = color,
         border_color = "black"
         )+coord_fixed(expand=2)
sero.color<-"#4f2e91ff"

```

## calculate # of eq.subclass / neurotransmitter type 
```{r}
MM.NT.NeqS<-data.frame(table(NT.all.mapped.HY$MM.NT))
colnames(MM.NT.NeqS)<-c("MM.NT","MM.NT.count")
sum<-sum(MM.NT.NeqS$MM.NT.count)
MM.NT.NeqS<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist", "Gly", "Sero-Dopa", "Sero-Hist"),
  eq.subclass.count=c(23/sum,26/sum, 10/sum, 1/sum, 9/sum, 0/sum, 0/sum, 1/sum, 0/sum, 0/sum, 0),
  species=rep("MM",11)
                        )
#MO
MO.nt<-read.csv(file="./MO.nt.type.csv")
MO.NT.NeqS<-data.frame(table(MO.nt$Neurotransmitter))
colnames(MO.NT.NeqS)<-c("MO.NT","MO.NT.count")
sum<-sum(MO.NT.NeqS$MO.NT.count)
MO.NT.NeqS<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist","Gly", "Sero-Dopa", "Sero-Hist","unclear"),
  eq.subclass.count=c(69/sum, 72/sum, 3/sum, 1/sum, 4/sum, 0/sum, 0/sum, 1/sum, 0/sum, 0/sum, 0, 13/sum),
  species=rep("MO",12)
)
  
# CJ
CJ.nt<-read.csv(file="./CJ.nt.type.csv")
CJ.NT.NeqS<-data.frame(table(CJ.nt$Neurotransmitter))
colnames(CJ.NT.NeqS)<-c("CJ.NT","CJ.NT.count")
sum<-sum(CJ.NT.NeqS$CJ.NT.count)
CJ.NT.NeqS<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist","Gly", "Sero-Dopa", "Sero-Hist","unclear"),
  eq.subclass.count=c(54/sum, 65/sum, 1/sum, 1/sum, 5/sum, 0, 0, 1/sum, 0, 0, 0, 16/sum),
  species=rep("CJ",12)
)

# AC
AC.nt<-read.csv(file="./AC.nt.type.csv")
AC.NT.NeqS<-data.frame(table(AC.nt$Neurotransmitter))
colnames(AC.NT.NeqS)<-c("AC.NT","AC.NT.count")
sum<-sum(AC.NT.NeqS$AC.NT.count)
AC.NT.NeqS<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist","Gly", "Sero-Dopa", "Sero-Hist","unclear"),
  eq.subclass.count=c(49/sum, 55/sum, 1/sum, 1/sum, 1/sum, 1/sum, 0, 0, 0, 0, 1/sum, 18/sum),
  species=rep("AC",12)
)

# XT
XT.nt<-read.csv(file="./XT.nt.type.csv")
XT.NT.NeqS<-data.frame(table(XT.nt$Neurotransmitter))
colnames(XT.NT.NeqS)<-c("XT.NT","XT.NT.count")
sum<-sum(XT.NT.NeqS$XT.NT.count)
XT.NT.NeqS<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist","Gly", "Sero-Dopa", "Sero-Hist","unclear"),
  eq.subclass.count=c(48/sum, 44/sum, 1/sum, 2/sum, 5/sum, 1/sum, 0, 1/sum, 9/sum, 0, 0, 9/sum),
  species=rep("XT",12)
)

# DR
DR.nt<-read.csv(file="./DR.nt.type.csv")
DR.NT.NeqS<-data.frame(table(DR.nt$Neurotransmitter))
colnames(DR.NT.NeqS)<-c("DR.NT","DR.NT.count")
sum<-sum(DR.NT.NeqS$DR.NT.count)
DR.NT.NeqS<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist","Gly", "Sero-Dopa", "Sero-Hist","unclear"),
  eq.subclass.count=c(21/sum, 31/sum, 2/sum, 1/sum, 2/sum, 0/sum, 0, 1/sum, 0, 2/sum, 0, 25/sum),
  species=rep("DR",12)
                        )

all.NT.NeqS <- rbind(MO.NT.NeqS, CJ.NT.NeqS)
all.NT.NeqS <- rbind(all.NT.NeqS, AC.NT.NeqS)
all.NT.NeqS <- rbind(all.NT.NeqS, XT.NT.NeqS)
all.NT.NeqS <- rbind(all.NT.NeqS, DR.NT.NeqS)

all.NT.NeqS <- within(all.NT.NeqS, 
                       species <- factor(species, levels=c("MO","CJ","AC","XT","DR"))
)

color=c("#73E785","#fcf04b","#FF3358","#2B93DF","#0a9964","#820e57","#ff7621","#03EDFF","#4f2e91ff", "#4f2e91ff", "#ff7621","#99999960")
eq.subclass.plot <- ggplot(all.NT.NeqS, aes(fill=NT.type, y=eq.subclass.count, x=species)) + 
  geom_bar(position="stack", stat="identity")+
  scale_fill_manual(values = color)+
  theme_classic()+
  ylab("%eq.Subclass")

```

## calculate # of cells / neurotransmitter type

```{r}
#MM
#extract AI HY subclass cell - NT table
MM.NT.Ncell <- data.frame(cell.id=rownames(metadata.AI.HYsubclass), NT.type = metadata.AI.HYsubclass$neurotransmitter, subclass=metadata.AI.HYsubclass$subclass)

#relabel
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="058 PAL-STR Gaba-Chol"] <- "GABA-Chol"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="081 ACB-BST-FS D1 Gaba"] <- "Glut-GABA"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="085 SI-MPO-LPO Lhx8 Gaba"] <- "Glut-GABA"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="089 PVR Six3 Sox3 Gaba"] <- "GABA-Dopa"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="092 TMv-PMv Tbx3 Hist-Gaba"] <- "GABA-Hist"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="098 AHN-SBPV-PVHd Pdrm12 Gaba"] <- "GABA-Dopa"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="099 SBPV-PVa Six6 Satb2 Gaba"] <- "GABA-Dopa"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="101 ZI Pax6 Gaba"] <- "GABA-Dopa"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="102 DMH-LHA Gsx1 Gaba"] <- "GABA-Chol"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="106 PVpo-VMPO-MPN Hmx2 Gaba"] <- "GABA-Dopa"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="107 DMH Hmx2 Gaba"] <- "Glut-GABA-Dopa"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="115 MS-SF Bsx Glut"] <- "Glut-GABA"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="119 SI-MA-LPO-LHA Skor1 Glut"] <- "Glut-GABA"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="125 DMH Hmx2 Glut"] <- "Glut-GABA"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="126 ARH-PVp Tbx3 Glut"] <- "Glut-GABA"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="131 LHA-AHN-PVH Otp Trh Glut"] <- "Glut-GABA"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="133 PVH-SO-PVa Otp Glut"] <- "Glut-GABA"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="138 PH Pitx2 Glut"] <- "Glut-Dopa"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="141 PH-SUM Foxa1 Glut"] <- "Glut-GABA"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="145 MH Tac2 Glut"] <- "Glut-Chol"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="159 IF-RL-CLI-PAG Foxa1 Glut"] <- "Glut-Dopa"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="210 PRT Mecom Gaba"] <- "GABA-Glyc"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="213 SCsg Gabrr2 Gaba"] <- "Glut-GABA"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="215 SNc-VTA-RAmb Foxa1 Dopa"] <- "Glut-GABA-Dopa"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="243 PGRN-PARN-MDRN Hoxb5 Glut"] <- "Glut-GABA"
MM.NT.Ncell$NT.type[MM.NT.Ncell$subclass=="285 MY Lhx1 Gly-Gaba"] <- "GABA-Glyc"
#unify convention
MM.NT.Ncell$NT.type[MM.NT.Ncell$NT.type=="GABA"]<-"Gaba"
MM.NT.Ncell$NT.type[MM.NT.Ncell$NT.type=="GABA-Chol"]<-"Gaba-Chol"
MM.NT.Ncell$NT.type[MM.NT.Ncell$NT.type=="Glut-GABA"]<-"Glut-Gaba"
MM.NT.Ncell$NT.type[MM.NT.Ncell$NT.type=="GABA-Dopa"]<-"Gaba-Dopa"
MM.NT.Ncell$NT.type[MM.NT.Ncell$NT.type=="GABA-Hist"]<-"Gaba-Hist"
MM.NT.Ncell$NT.type[MM.NT.Ncell$NT.type=="Glut-GABA-Dopa"]<-"Glut-Gaba-Dopa"
MM.NT.Ncell$NT.type[MM.NT.Ncell$NT.type=="GABA-Glyc"]<-"Gaba-Gly"

table.MM <- table(MM.NT.Ncell$NT.type)
sum<-sum(table.MM)
table.MM <- table.MM/sum
MM.NT.Ncell<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist","Gly"),
  cell.count=c(sum(table.MM[c('Glut','Glut-Dopa')]), sum(table.MM[c('Gaba','Gaba-Chol','Gaba-Dopa','Gaba-Hist')]),sum(table.MM[c('Glut-Gaba','Glut-Gaba-Dopa')]), sum(table.MM[c('Gaba-Chol')]), sum(table.MM[c('Gaba-Dopa','Glut-Dopa','Glut-Gaba-Dopa')]), 0, 0, sum(table.MM[c('Gaba-Hist')]), 0),
  species=rep("MM",9)
  )

## MO
table.MO <- table(scdata.MO$neurotransmitter)
sum<-sum(table.MO)
table.MO <- table.MO/sum
MO.NT.Ncell<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist","Gly", "Sero-Dopa", "Sero-Hist","unclear"),
  cell.count=c(sum(table.MO[c('Glut')]), sum(table.MO[c('Gaba')]), sum(table.MO[c('Glut-Gaba')], na.rm = TRUE), sum(table.MO[c('Chol','Glut-Chol','Gaba-Chol')], na.rm = TRUE), sum(table.MO[c('Gaba-Dopa','Glut-Dopa','Glut-Gaba-Dopa')], na.rm = TRUE), sum(table.MO[c('Gaba-Sero','Glut-Sero','Sero')], na.rm = TRUE), sum(table.MO[c('Nora')], na.rm = TRUE), sum(table.MO[c('Gaba-Hist','Glut-Hist','Gaba-Glut-Hist','Hist')], na.rm = TRUE), sum(table.MO[c('Gaba-Gly','Glut-Gly','Gaba-Glut-Gly','Gly')], na.rm = TRUE), sum(table.MO[c('Gaba-Dopa-Sert')], na.rm = TRUE), sum(table.MO[c('Glut-Gaba-Sert-Hist')], na.rm = TRUE), sum(table.MO[c('unclear')])),
  species=rep("MO",12)
  ) 
  
## CJ
table.CJ <- table(scdata.CJ$neurotransmitter)
sum<-sum(table.CJ)
table.CJ <- table.CJ/sum
CJ.NT.Ncell<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist","Gly", "Sero-Dopa", "Sero-Hist","unclear"),
  cell.count=c(sum(table.CJ[c('Glut')]), sum(table.CJ[c('Gaba')]), sum(table.CJ[c('Glut-Gaba')], na.rm = TRUE), sum(table.CJ[c('Chol','Glut-Chol','Gaba-Chol')], na.rm = TRUE), sum(table.CJ[c('Gaba-Dopa','Glut-Dopa','Glut-Gaba-Dopa')], na.rm = TRUE), sum(table.CJ[c('Gaba-Sero','Glut-Sero','Sero')], na.rm = TRUE), sum(table.CJ[c('Nora')], na.rm = TRUE), sum(table.CJ[c('Gaba-Hist','Glut-Hist','Gaba-Glut-Hist','Hist')], na.rm = TRUE), sum(table.CJ[c('Gaba-Gly','Glut-Gly','Gaba-Glut-Gly','Gly')], na.rm = TRUE), sum(table.CJ[c('Gaba-Dopa-Sert')], na.rm = TRUE), sum(table.CJ[c('Glut-Gaba-Sert-Hist')], na.rm = TRUE), sum(table.CJ[c('unclear')])),
  species=rep("CJ",12)
  ) 

## AC
table.AC <- table(scdata.AC$neurotransmitter)
sum<-sum(table.AC)
table.AC <- table.AC/sum
AC.NT.Ncell<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist","Gly", "Sero-Dopa", "Sero-Hist","unclear"),
  cell.count=c(sum(table.AC[c('Glut')]), sum(table.AC[c('Gaba')]), sum(table.AC[c('Glut-Gaba')], na.rm = TRUE), sum(table.AC[c('Chol','Glut-Chol','Gaba-Chol')], na.rm = TRUE), sum(table.AC[c('Gaba-Dopa','Glut-Dopa','Glut-Gaba-Dopa')], na.rm = TRUE), sum(table.AC[c('Gaba-Sero','Glut-Sero','Sero')], na.rm = TRUE), sum(table.AC[c('Nora')], na.rm = TRUE), sum(table.AC[c('Gaba-Hist','Glut-Hist','Gaba-Glut-Hist','Hist')], na.rm = TRUE), sum(table.AC[c('Gaba-Gly','Glut-Gly','Gaba-Glut-Gly','Gly')], na.rm = TRUE), sum(table.AC[c('Gaba-Dopa-Sert')], na.rm = TRUE), sum(table.AC[c('Glut-Gaba-Sert-Hist')], na.rm = TRUE), sum(table.AC[c('unclear')])),
  species=rep("AC",12)
  ) 

## XT
table.XT <- table(scdata.XT$Neurotransmitter)
sum<-sum(table.XT)
table.XT <- table.XT/sum
XT.NT.Ncell<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist","Gly", "Sero-Dopa", "Sero-Hist","unclear"),
  cell.count=c(sum(table.XT[c('Glut')]), sum(table.XT[c('Gaba')]), sum(table.XT[c('Glut-Gaba')], na.rm = TRUE), sum(table.XT[c('Chol','Glut-Chol','Gaba-Chol')], na.rm = TRUE), sum(table.XT[c('Gaba-Dopa','Glut-Dopa','Glut-Gaba-Dopa')], na.rm = TRUE), sum(table.XT[c('Gaba-Sero','Glut-Sero','Sero')], na.rm = TRUE), sum(table.XT[c('Nora')], na.rm = TRUE), sum(table.XT[c('Gaba-Hist','Glut-Hist','Gaba-Glut-Hist','Hist')], na.rm = TRUE), sum(table.XT[c('Gaba-Gly','Glut-Gly','Gaba-Glut-Gly','Gly')], na.rm = TRUE), sum(table.XT[c('Gaba-Dopa-Sert')], na.rm = TRUE), sum(table.XT[c('Glut-Gaba-Sert-Hist')], na.rm = TRUE), sum(table.XT[c('unclear')])),
  species=rep("XT",12)
  ) 


## DR
table.DR <- table(scdata.DR$neurotransmitter)
sum<-sum(table.DR)
table.DR <- table.DR/sum
DR.NT.Ncell<-data.frame(
  NT.type=c("Glut","Gaba","Glut-Gaba","Chol","Dopa","Sero","Nora","Hist","Gly", "Sero-Dopa", "Sero-Hist","unclear"),
  cell.count=c(sum(table.DR[c('Glut')]), sum(table.DR[c('Gaba')]), sum(table.DR[c('Glut-Gaba')], na.rm = TRUE), sum(table.DR[c('Chol','Glut-Chol','Gaba-Chol')], na.rm = TRUE), sum(table.DR[c('Gaba-Dopa','Glut-Dopa','Glut-Gaba-Dopa')], na.rm = TRUE), sum(table.DR[c('Gaba-Sero','Glut-Sero','Sero')], na.rm = TRUE), sum(table.DR[c('Nora')], na.rm = TRUE), sum(table.DR[c('Gaba-Hist','Glut-Hist','Gaba-Glut-Hist','Hist')], na.rm = TRUE), sum(table.DR[c('Gaba-Gly','Glut-Gly','Gaba-Glut-Gly','Gly')], na.rm = TRUE), sum(table.DR[c('Gaba-Dopa-Sert')], na.rm = TRUE), sum(table.DR[c('Glut-Gaba-Sert-Hist')], na.rm = TRUE), sum(table.DR[c('unclear')])),
  species=rep("DR",12)
  ) 



all.NT.Ncell <- rbind(MO.NT.Ncell, CJ.NT.Ncell)
all.NT.Ncell <- rbind(all.NT.Ncell, AC.NT.Ncell)
all.NT.Ncell <- rbind(all.NT.Ncell, XT.NT.Ncell)
all.NT.Ncell <- rbind(all.NT.Ncell, DR.NT.Ncell)

all.NT.Ncell <- within(all.NT.Ncell, 
                       species <- factor(species, levels=c("MO","CJ","AC","XT","DR"))
                       )

color=c("#73E785","#fcf04b","#FF3358","#2B93DF","#0a9964","#820e57","#ff7621","#03EDFF","#4F2E91FF","#4F2E9160","#ff762160","#99999960")
Ncell.NT.plot <- ggplot(all.NT.Ncell, aes(fill=NT.type, y=cell.count, x=species)) + 
  geom_bar(position="stack", stat="identity")+
  scale_fill_manual(values = color)+
  theme_classic()+
  ylab("% Cell/NT type")


  
```
